### 2010-03-23 paper.tex
Add paper template for CCP.

\documentclass[mathpazo]{cicp}

%%%%% author macros %%%%%%%%%
% place your own macros HERE
%%%%% end %%%%%%%%%

\begin{document}
%%%%% title : short title may not be used but TITLE is required.
% \title{TITLE}
% \title[short title]{TITLE}
\title{Here is the Title}

%%%%% author(s) :
% single author:
% \author[name in running head]{AUTHOR\corrauth}
% [name in running head] is NOT OPTIONAL, it is a MUST.
% Use \corrauth to indicate the corresponding author.
% Use \email to provide email address of author.
% \footnote and \thanks are not used in the heading section.
% Another acknowlegments/support of grants, state in Acknowledgments section
% \section*{Acknowledgments}
\author[O.~Author]{Only Author\corrauth}
\address{School of Mathematical Sciences, Beijing Normal University,
Beijing 100875, P.R. China}
\email{{\tt author@email} (O.~Author)}

% multiple authors:
% Note the use of \affil and \affilnum to link names and addresses.
% The author for correspondence is marked by \corrauth.
% use \emails to provide email addresses of authors
% e.g. below example has 3 authors, first author is also the corresponding
%      author, author 1 and 3 having the same address.
% \author[Zhang Z R et.~al.]{Zhengru Zhang\affil{1}\comma\corrauth,
%       Author Chan\affil{2}, and Author Zhao\affil{1}}
% \address{\affilnum{1}\ School of Mathematical Sciences,
%          Beijing Normal University,
%          Beijing 100875, P.R. China. \\
%           \affilnum{2}\ Department of Mathematics,
%           Hong Kong Baptist University, Hong Kong SAR}
% \emails{{\tt zhang@email} (Z.~Zhang), {\tt chan@email} (A.~Chan),
%          {\tt zhao@email} (A.~Zhao)}
% \footnote and \thanks are not used in the heading section.
% Another acknowlegments/support of grants, state in Acknowledgments section
% \section*{Acknowledgments}

%%%%% Begin Abstract %%%%%%%%%%%
\begin{abstract}
The abstract should provide a brief summary of the main findings of the paper.
\end{abstract}
%%%%% end %%%%%%%%%%%

%%%%% AMS/PACs/Keywords %%%%%%%%%%%
%\pac{}
\ams{52B10, 65D18, 68U05, 68U07}
\keywords{moving mesh method, conservative interpolation, iterative method, $l^2$ projection.}

%%%% maketitle %%%%%
\maketitle

%%%% Start %%%%%%
\section{Introduction}
\label{sec1}
In the past two decades, there has been important progress in developing adaptive mesh methods for PDEs.
Mesh adaptivity is usually of two types in form: local mesh refinement and moving mesh method.  ....

\section{Preparation of Manuscript}
\label{sec2}
The Title Page should contain the article title, authors' names and complete affiliations,
and email addresses of all authors. The Abstract should provide a brief summary of the main findings of the paper.

References should be cited in the text by a number in square brackets.
Literature cited should appear on a separate page at the end of the article
and should be styled and punctuated using standard abbreviations for journals
(see Chemical Abstracts Service Source Index, 1989). For unpublished lectures of symposia,
include title of paper, name of sponsoring society in full, and date.
Give titles of unpublished reports with "(unpublished)" following the reference.
Only articles that have been published or are in press should be included in the references.
Unpublished results or personal communications should be cited as such in the text.
Please note the sample at the end of this paper.

Equations should be typewritten whenever possible and the number placed in parentheses at the right margin.
Reference to equations should use the form "Eq. (2.1)" or simply "(2.1)." Superscripts and subscripts should
be typed or handwritten clearly above and below the line, respectively.

Figures should be in a finished form suitable for publication. Number figures consecutively with Arabic numerals.
Lettering on drawings should be of professional quality or generated by high-resolution computer graphics and must be
large enough to withstand appropriate reduction for publication.
For example, if you use {\sf MATLAB} to do figure plots,
axis labels should be at least point 18. Title should be 24 points or above. Tick marks labels
better have 14 points or above. Line width should be 2 (or above).

Illustrations in color in most cases can be accepted
only if the authors defray the cost. At the Editor's discretion a limited number of color figures each year of special
interest will be published at no cost to the author.

%%%% Acknowledgments %%%%%%%%
\section*{Acknowledgments}
The author would like to thank  ....

%%%% Bibliography  %%%%%%%%%%
\begin{thebibliography}{99}
\bibitem{Berger}M. J. Berger and P. Collela, Local adaptive mesh refinement
for shock hydrodynamics,
J. Comput. Phys., 82 (1989), 62-84.
\bibitem{deBoor}C. de Boor,  Good Approximation By Splines With Variable Knots II, in Springer Lecture
 Notes Series 363, Springer-Verlag, Berlin, 1973.
\bibitem{TanTZ} Z. J. Tan, T. Tang and Z. R. Zhang, A simple moving mesh method for one- and
two-dimensional phase-field equations, J. Comput. Appl. Math., to appear.
\bibitem{Toro}E. F. Toro, Riemann Solvers and Numerical Methods for Fluid Dynamics,
Springer-Verlag Berlin Heidelbert, 1999.
\end{thebibliography}

\end{document}

### 2010-04-01 paper.tex
Revised introduction. Needs more work on it. Will continue with the LBM.

@@
 \documentclass[mathpazo]{cicp}
+\usepackage{bm}
+\usepackage{graphicx}
+\usepackage{color}
+%\usepackage[sort]{cite}
+%\usepackage[sort&compress,authoryear,round]{natbib}
 
@@
 % \title[short title]{TITLE}
-\title{Here is the Title}
+\title{The effecient GPU lattice Boltzmann multicomponent and multiphase systems modelling}
 
@@
 % \section*{Acknowledgments}
-\author[O.~Author]{Only Author\corrauth}
-\address{School of Mathematical Sciences, Beijing Normal University,
-Beijing 100875, P.R. China}
-\email{{\tt author@email} (O.~Author)}
+\author[M.~Januszewski and A.~Kuzmin]{M.~Januszewski\affil{1}\corrauth and A.~Kuzmin\affil{2}}
+\address{\affilnum{1}\ University, Poland\\
+\affilnum{2}\ Department of Mechanical and Manufacturing Engineering,
+Schulich School of Engineering,
+University of Calgary, Canada
+}
+\emails{{\tt michalj@gmail.com} (M.~Januszewski),{\tt shurik.kuzmin@gmail.com} (A.~Kuzmin)}
 
@@
 \begin{abstract}
-The abstract should provide a brief summary of the main findings of the paper.
+Despite the number and successful implementations of number of lattice Boltzmann systems on Graphical Processor Units (GPUs), the implementation of more complicated problems as multiphase models or Nearest-Neighbours (NN) interaction, which includes all the boundary conditions, on GPUs is still not well developed.
+The purpose of the paper is twofold: to develop the efficient GPUs implementation algorithms for the binary liquids flows, which are subclass of the NN interactions, and to illustrate the usefullness in terms of performance of GPUs for multiphase models simulations.
 \end{abstract}
@@
 %\pac{}
-\ams{52B10, 65D18, 68U05, 68U07}
-\keywords{moving mesh method, conservative interpolation, iterative method, $l^2$ projection.}
+
+\ams{52B10, 65D18, 68U05, 68U07{\color{red} We need to find somewhere classifications.}}
+\keywords{Lattice Boltzmann method, GPU, Shan-Chen model, Free-energy model, Binary liquids}
 
@@
 \section{Introduction}
-\label{sec1}
-In the past two decades, there has been important progress in developing adaptive mesh methods for PDEs.
-Mesh adaptivity is usually of two types in form: local mesh refinement and moving mesh method.  ....
+\label{sec:introduction}
+During the last two decades since the lattice Boltzmann method was introduced \cite{mcnamara}, the method became mature to compete with the well established computational fluid dynamics (CFD).
+Due to the kinetic nature of the method the method is applicable to simulate not only hydrodynamics but also thermal flows \cite{karlin-minimalmodels,yuan-thermal}, micro-flows \cite{ansumali-small-knudsen}, ferrofluids \cite{rosensweig,kuzmin-aniso}, multiphase flows \cite{rothman-color,Shan-chen:extended,swift}.
+In the presence of complex physical effects such as turbulence, multi-phase flows, flow of oil and gas through porous material, flow around surfaces with complex geometries, or blood flow with deformable particles, the solution of the basic Navier-Stokes equations becomes difficult or impractical.
+On the other hand, the micro-scale calculations for such systems based on the  Molecular Dynamics approach (MD) are computationally extremely demanding.
+The Lattice Boltzmann Method (LBM), which focuses on the meso-scale, may be the best choice for such complex flow problems, being able to capture macro- and micro-flows.
 
-\section{Preparation of Manuscript}
-\label{sec2}
-The Title Page should contain the article title, authors' names and complete affiliations,
-and email addresses of all authors. The Abstract should provide a brief summary of the main findings of the paper.
+{\color{red} Michal, please add here more details.}
+The lattice Boltzmann method is the explicit numerical method opposing some restrictions on the time propagation step.
+Thus the physical time step corresponding to the numerical time step is relatively small.
+Performing large-time-scale simulations become unpractical due to the high computational demand.
+To overcome the problem of the high computational demand the physical problems are calculated with the help of parallel processors of a cluster.
+The recent years trend is to perform the simulations on Graphical Processing Units (GPUs) which can be viewed as small supercomputers due to the large number of independent threads performing a simulation.
+The attractive part of simulations on GPUs is the low price (in comparison with clusters) and large performance.
+Not all the problems are matched to be simulated on parallel machines.
+However, the lattice Boltzmann method is a perfect match for parallel structure machines as GPUs.
 
-References should be cited in the text by a number in square brackets.
-Literature cited should appear on a separate page at the end of the article
-and should be styled and punctuated using standard abbreviations for journals
-(see Chemical Abstracts Service Source Index, 1989). For unpublished lectures of symposia,
-include title of paper, name of sponsoring society in full, and date.
-Give titles of unpublished reports with "(unpublished)" following the reference.
-Only articles that have been published or are in press should be included in the references.
-Unpublished results or personal communications should be cited as such in the text.
-Please note the sample at the end of this paper.
+During the last years the method was applied for many different problems, such as, to name a few, the calculation of the drag coefficient for the two-dimensional flow through cylinders \cite{tolke-twod} and moving sphere in three-dimensional channel \cite{tolke-GPU}.
+To the authors' best knowledge only simple hydrodynamics problems GPUs implementation is thoroughly described in the literature.
+However, the nearest neighbours (NN) interaction which includes advanced boundary lattice Boltzmann methods and multiphase problems are still not covered.
+For many multiphase problems the local density gradient is required to properly simulate the physics.
+The implementation of the multiphase models are more challenging than simple hydrodynamics problems because of the additional information required from other nodes.
+The simple hydrodynamics lattice Boltzmann method is local and is perfectly parallelized.
+The NN problems requires the additional information from other nodes and can't be addressed as local any more.
 
-Equations should be typewritten whenever possible and the number placed in parentheses at the right margin.
-Reference to equations should use the form "Eq. (2.1)" or simply "(2.1)." Superscripts and subscripts should
-be typed or handwritten clearly above and below the line, respectively.
+This work's aim is to present the efficient GPUs algorithms to implement the binary liquids models based on the free-energy approach and the Shan-Chen method.
+First, we thoroughly describe the GPUs algorithms.
+Then the Shan-Chen model and the free-energy binary liquid model lattice Boltzmann implementation is presented.
+The numerical results for the comparison benchmark between those two models is done in terms of performance and accuracy in the ``Numerical results`` section.
+We conclude the paper with the main findings.
 
-Figures should be in a finished form suitable for publication. Number figures consecutively with Arabic numerals.
-Lettering on drawings should be of professional quality or generated by high-resolution computer graphics and must be
-large enough to withstand appropriate reduction for publication.
-For example, if you use {\sf MATLAB} to do figure plots,
-axis labels should be at least point 18. Title should be 24 points or above. Tick marks labels
-better have 14 points or above. Line width should be 2 (or above).
+\section{LBM}
+\label{sec:lbm:binary:liquids}
+As mentioned previously, one of the most popular multiphase models is the Shan-Chen model.
+Originally introduced in 1993 \cite{Shan-chen:extended} the model attracted a multitude of researchers by its simplicity.
+The results of the model were applied in simulation of two-phase blood flow \cite{halliday-multicomponent}, flow in microchannels \cite{fan-multi}, thermal flow \cite{zhang-thermal}, and studies of break-up of the liquid droplet \cite{nourgaliev-breakup}.
+Many researchers use the model in its original formulation, which has many discrepancies with physics, such as the viscosity dependance of steady-state droplet condition.
+Thus, the aim of this thesis is to explore the basics of the Shan-Chen model and to resolve the problems associated with it.
+In this chapter the theory behind the original Shan-Chen model will be presented.
+The problems coming from the formulation will be enumerated, i.e.
+the achieved gas-liquid ratio limit, the non-thermodynamical behaviour, the dependence of the surface tension on the same parameter as the equation of state.
+All of the limitations will be thoroughly examined and supported simulations.
+Some limitations of the Shan-Chen model will be overcome.
+For instance, the stability limit can be expanded by applying the extended equilibrium distribution function.
+The dependence of the surface tension and the equation of state on one parameter is solved with the multirange potential.
+Also, the extension of the Shan-Chen model, the so-called generalized Shan-Chen model, is presented in order to simulate ferro-fluidics.
 
-Illustrations in color in most cases can be accepted
-only if the authors defray the cost. At the Editor's discretion a limited number of color figures each year of special
-interest will be published at no cost to the author.
+For simplicity, the two-dimensional implementation of the Shan-Chen model with 9 velocities, the D2Q9 model, is used in this chapter.
+However, the results are general in nature and can be applied to three-dimensional problems.
 
+\subsection{Shan-Chen model}
+
+In the Shan-Chen model \cite{Shan-chen:extended} the force at a given node depends on all local neighbours characteristics:
+\begin{equation}  \label{Shan-Chen:Shan-Chen:cont}
+\bm{F}=G\psi_{0}\sum_{i}{w_i \psi_{i}\bm{c_{i}}},
+\end{equation}
+where $\psi$ is the function of local characteristics of node, i.e.
+density.
+The function $\psi$ is usually in the literature noted as the pseudopotential.
+In the above, $\psi_0$ denotes the local value at a given lattice site, and the summation, $i$, is taken over
+neighbours of the node defined by the discrete velocity set $\{\bm{c_i}\}$.
+Coefficient $G$ is responsible for the attraction force between the molecules.
+Overall, the Shan-Chen force is the approximation of the force between molecules \cite{kwok,kwok-contact-angle}.
+
+\subsection{Equation of state}
+The equation of state for the LBE \cite{Succi-book} is $P=c_s^2\rho$, as it was shown in equation (\ref{navier-stokes:with:everything}), where $c_s$ is the scaling parameter which for the D2Q9 model equals $c_s=1/\sqrt{3}$.
+However, the inclusion of a density-dependent force term brings about a non-ideal contribution.
+Using the Taylor expansion, the Shan-Chen force (\ref{Shan-Chen:Shan-Chen:cont}) for the D2Q9 model can be expressed as follows:
+\begin{equation}
+\begin{aligned}
+\bm{F}=&G\psi(x,y)(w_1 \psi(x+c\Delta t,y)\bm{c_{1}}+w_2 \psi(x,y+c\Delta t)\bm{c_{2}}%
++w_3 \psi(x-c\Delta t,y)\bm{c_{3}}+ \\
+&w_4 \psi(x,y-c\Delta t)\bm{c_{4}}+w_5 \psi(x+c\Delta t,y+c\Delta t)\bm{c_{5}}+ \\
+&w_6 \psi(x-c\Delta t,y+c\Delta t)\bm{c_{6}}+w_7 \psi(x-c\Delta t,y-c\Delta t)\bm{%
+c_{7}}+ \\
+&w_8 \psi(x+c\Delta t,y-c\Delta t)\bm{c_{8}})=G\psi(\frac{1}{3}\bm{\nabla}\psi+%
+\frac{1}{18}\bm{\nabla}\Delta\psi).
+\end{aligned}
+\end{equation}
+The following formulation is derived algebraically:
+\begin{equation}
+\label{shan-chen:force:taylor:expansion}
+\begin{aligned}
+F_{\alpha}&=G\psi\left(\frac{1}{3}\partial_{\alpha}\psi+\frac{1}{18}\partial_{\alpha}\Delta%
+\psi\right) =G\left(\frac{1}{6}\partial_{\alpha}\psi^2+\frac{1}{18}%
+\biggl(\partial_{\alpha}(\psi\Delta\psi)-\Delta\psi\partial_{\alpha}\psi\biggr)\right) \\
+&=G\left(\frac{1}{6}\partial_{\alpha}\psi^2+\frac{1}{18}%
+\biggl(\partial_{\alpha}(\psi\Delta\psi)+\frac{1}{2}\partial_{\alpha}|\bm{\nabla}\psi|^2
+-\partial_{\beta}\partial_{\alpha}\psi\partial_{\beta}\psi\biggr)\right).
+\end{aligned}
+\end{equation}
+The force influence can be included to the momentum-flux tensor \cite{Sbragaglia-Succi}:
+\begin{equation*}
+\partial_{\beta}P_{\alpha\beta}=-F_{\alpha}+\partial_{\alpha}p=-F_{\alpha}+\partial_{\alpha}(c_s^2%
+\rho).
+\end{equation*}
+Thus, the flux tensor $P_{\alpha\beta}$ is modified as follows:
+\begin{equation*}
+P_{\alpha\beta}=\biggl(c_s^2\rho+\frac{G}{6}\psi^2+\frac{G}{36}%
+|\bm{\nabla}\psi|^2 +\frac{G}{18}\psi\Delta\psi\biggr)\delta_{\alpha\beta}-\frac{G}{%
+18}\partial_{\alpha}\psi\partial_{\beta}\psi.
+\end{equation*}
+By analogy with classical mechanics, the potential of the force can be introduced as:
+\begin{equation}
+\label{force:potential}
+U=\frac{G}{6}\psi^2+\frac{G}{36}|\bm{\nabla}\psi|^2+\frac{G}{18}%
+\psi\Delta\psi.
+\end{equation}
+
+Since the gradient terms in equation (\ref{force:potential}) are small compared with the leading terms (the characteristic length of the interface is longer than the lattice spacing, as in
+all diffuse-interface methods), the equation (\ref{force:potential}) can be approximated as
+\begin{equation}
+p=\rho c_s^2+\frac{G}{6}\psi^2.
+\end{equation}
+
+\subsection{Binary liquid model}
+\section{Binary liquid multiphase model}
+In this section we'd like to describe binary liquid model. Binary model is based on free energy model. The free energy for binary liquids can be written as \cite{swift,landau}:
+\begin{equation}
+F=\int{J \mathrm{d}V}=\int{\left(c_s^2\rho\ln\rho+\frac{A}{2}\phi^2+\frac{B}{4}\phi^4 + \frac{k}{2}(\nabla \phi)^2 \right)\mathrm{d}V} \quad,
+\end{equation}
+where coefficients $A$ and $B$ are responsible for the phase separation, coefficient $k$ is responsible for the surface tension between phases and $\phi$ is the order parameter.
+Notice that if $A<0$ then the phase separation occurs, and $\phi$ will take values as $-1$ and $1$ for equilibrium phases.
+
+By minimising free energy functional with constraints for constant number of particles, one can obtain the chemical potential:
+\begin{equation}
+\mu=A\phi+B\phi^3 - k \Delta \phi
+\end{equation}
+From free energy model we can construct the pressure tensor:
+\begin{equation}
+P_{\alpha \beta}=\bigg(p_0 - k\phi\Delta\phi-k/2 |\nabla \phi|^2\bigg)\delta_{\alpha\beta}+k\partial_{\alpha}\phi\partial_{\beta}\phi \quad,
+\end{equation}
+where
+\begin{equation}
+ p_0=\rho c_s^2+\frac{A}{2}\phi^2+\frac{3}{4}B \phi^2
+\label{binary:bulk:pressure}
+\end{equation}
+
+Let us take parameter $A$ as $-B$, then after substitution we will obtain the following bulk pressure:
+\begin{equation}
+ p_0=\rho c_s^2-\frac{B}{2}\phi^2+\frac{3}{4}B \phi^2
+\end{equation}
+
+The whole system of equations which describe binary model is the following:
+\begin{equation}
+\begin{aligned}
+&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
+&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=\\
+&-\partial_{\alpha}P_{\alpha \beta} + \nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha} + \frac{1}{3}\partial_{\gamma} u_{\gamma} \delta_{\alpha\beta}\right)\\
+&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=D\Delta \mu
+\end{aligned}
+\label{binary:fluid:system}
+\end{equation}
+
+We take equilibrium distribution function for the binary-liquid gas model. In general, it can be presented in the following form:
+\begin{equation}
+\begin{aligned}
+&g_i^{eq}=w_i(\Gamma \mu + \frac{\phi^m c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi^m \frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}), i\neq0 \\
+&g_0^{eq}=\phi^m-\sum_{i\neq0}{g_i^{eq}}\quad,
+\end{aligned}
+\end{equation}
+where $\mu$ is the chemical potential defined above, and $\Gamma$ is the mobility parameter.
+
+\section{GPU implementation}
+\section{Results}
+\section{Conclusion}
 %%%% Acknowledgments %%%%%%%%
 \section*{Acknowledgments}
-The author would like to thank  ....
+A.~Kuzmin wants to thank the Alberta Ingenuity Fund for their financial support.
 
 %%%% Bibliography  %%%%%%%%%%
-\begin{thebibliography}{99}
-\bibitem{Berger}M. J. Berger and P. Collela, Local adaptive mesh refinement
-for shock hydrodynamics,
-J. Comput. Phys., 82 (1989), 62-84.
-\bibitem{deBoor}C. de Boor,  Good Approximation By Splines With Variable Knots II, in Springer Lecture
- Notes Series 363, Springer-Verlag, Berlin, 1973.
-\bibitem{TanTZ} Z. J. Tan, T. Tang and Z. R. Zhang, A simple moving mesh method for one- and
-two-dimensional phase-field equations, J. Comput. Appl. Math., to appear.
-\bibitem{Toro}E. F. Toro, Riemann Solvers and Numerical Methods for Fluid Dynamics,
-Springer-Verlag Berlin Heidelbert, 1999.
-\end{thebibliography}
+\bibliographystyle{plainnat}
+\bibliography{paper}
+
+% \begin{thebibliography}{99}
+% \bibitem{Berger}M. J. Berger and P. Collela, Local adaptive mesh refinement
+% for shock hydrodynamics,
+% J. Comput. Phys., 82 (1989), 62-84.
+% \bibitem{deBoor}C. de Boor,  Good Approximation By Splines With Variable Knots II, in Springer Lecture
+%  Notes Series 363, Springer-Verlag, Berlin, 1973.
+% \bibitem{TanTZ} Z. J. Tan, T. Tang and Z. R. Zhang, A simple moving mesh method for one- and
+% two-dimensional phase-field equations, J. Comput. Appl. Math., to appear.
+% \bibitem{Toro}E. F. Toro, Riemann Solvers and Numerical Methods for Fluid Dynamics,
+% Springer-Verlag Berlin Heidelbert, 1999.
+% \end{thebibliography}
 

### 2010-04-01 paper.tex
Update my affiliation.
Add zipcode, fix GPUs in abstract, change the paper title.

@@
 % \title[short title]{TITLE}
-\title{The effecient GPU lattice Boltzmann multicomponent and multiphase systems modelling}
+\title{Efficient GPU implementation of multicomponent and multiphase Lattice Boltzmann models.}
 
@@
 \author[M.~Januszewski and A.~Kuzmin]{M.~Januszewski\affil{1}\corrauth and A.~Kuzmin\affil{2}}
-\address{\affilnum{1}\ University, Poland\\
+\address{\affilnum{1}\ Institute of Physics, University of Silesia, 40-007 Katowice, Poland\\
 \affilnum{2}\ Department of Mechanical and Manufacturing Engineering,
@@
 \begin{abstract}
-Despite the number and successful implementations of number of lattice Boltzmann systems on Graphical Processor Units (GPUs), the implementation of more complicated problems as multiphase models or Nearest-Neighbours (NN) interaction, which includes all the boundary conditions, on GPUs is still not well developed.
+Despite the number and successful implementations of number of lattice Boltzmann systems on Graphics Processing Units (GPUs), the implementation of more complicated problems as multiphase models or Nearest-Neighbours (NN) interaction, which includes all the boundary conditions, on GPUs is still not well developed.
 The purpose of the paper is twofold: to develop the efficient GPUs implementation algorithms for the binary liquids flows, which are subclass of the NN interactions, and to illustrate the usefullness in terms of performance of GPUs for multiphase models simulations.

### 2010-04-02 paper.tex
Changed to the numbering citations style. Natbib still doesn\'t work, but it\'s not required anymore.

@@
-\documentclass[mathpazo]{cicp}
+\documentclass[mathpazo,sort,numbers]{cicp}
 \usepackage{bm}
@@
 Schulich School of Engineering,
-University of Calgary, Canada
+University of Calgary, 2500 University Drive NW ,Calgary, Alberta, T2N 1N4 Canada
 }
@@
 %%%% Bibliography  %%%%%%%%%%
-\bibliographystyle{plainnat}
+\bibliographystyle{plain}
 \bibliography{paper}

### 2010-04-02 paper.tex
Cleaned the Shan-Chen and binary liquid implementations. Need to add binary Shan-Chen and one-component Julia's model. Probably to concentrate only on the one or binary components models.

@@
 \label{sec:lbm:binary:liquids}
-As mentioned previously, one of the most popular multiphase models is the Shan-Chen model.
+This works concentrates on two popular LBM multiphase models.
 Originally introduced in 1993 \cite{Shan-chen:extended} the model attracted a multitude of researchers by its simplicity.
-The results of the model were applied in simulation of two-phase blood flow \cite{halliday-multicomponent}, flow in microchannels \cite{fan-multi}, thermal flow \cite{zhang-thermal}, and studies of break-up of the liquid droplet \cite{nourgaliev-breakup}.
-Many researchers use the model in its original formulation, which has many discrepancies with physics, such as the viscosity dependance of steady-state droplet condition.
-Thus, the aim of this thesis is to explore the basics of the Shan-Chen model and to resolve the problems associated with it.
-In this chapter the theory behind the original Shan-Chen model will be presented.
-The problems coming from the formulation will be enumerated, i.e.
-the achieved gas-liquid ratio limit, the non-thermodynamical behaviour, the dependence of the surface tension on the same parameter as the equation of state.
-All of the limitations will be thoroughly examined and supported simulations.
-Some limitations of the Shan-Chen model will be overcome.
-For instance, the stability limit can be expanded by applying the extended equilibrium distribution function.
-The dependence of the surface tension and the equation of state on one parameter is solved with the multirange potential.
-Also, the extension of the Shan-Chen model, the so-called generalized Shan-Chen model, is presented in order to simulate ferro-fluidics.
-
-For simplicity, the two-dimensional implementation of the Shan-Chen model with 9 velocities, the D2Q9 model, is used in this chapter.
-However, the results are general in nature and can be applied to three-dimensional problems.
+The free-energy model \cite{swift} is based on the free-energy phase transition \cite{landau}.
+Both models werre successfully applied in simulation of two-phase blood flow \cite{halliday-multicomponent}, flow in microchannels \cite{pooley-contact}, thermal flow \cite{zhang-thermal}, and studies of break-up of the liquid droplet \cite{nourgaliev-breakup}.
+The section describes the two-dimensional implementation of the Shan-Chen and free-energy models with the D2Q9 model.
 
 \subsection{Shan-Chen model}
-
 In the Shan-Chen model \cite{Shan-chen:extended} the force at a given node depends on all local neighbours characteristics:
 \begin{equation}  \label{Shan-Chen:Shan-Chen:cont}
-\bm{F}=G\psi_{0}\sum_{i}{w_i \psi_{i}\bm{c_{i}}},
+\bm{F}(\bm{x})=-G\psi(\rho(\bm{x}))\sum_{i}{w_i \psi(\rho(\bm{x}+\bm{c_i}))\bm{c_{i}}},
 \end{equation}
-where $\psi$ is the function of local characteristics of node, i.e.
-density.
-The function $\psi$ is usually in the literature noted as the pseudopotential.
-In the above, $\psi_0$ denotes the local value at a given lattice site, and the summation, $i$, is taken over
-neighbours of the node defined by the discrete velocity set $\{\bm{c_i}\}$.
+where $\psi$ is the function of local characteristics of a node, and taken as $\psi(\rho)=1-\exp(-\rho)$.
 Coefficient $G$ is responsible for the attraction force between the molecules.
 Overall, the Shan-Chen force is the approximation of the force between molecules \cite{kwok,kwok-contact-angle}.
-
-\subsection{Equation of state}
-The equation of state for the LBE \cite{Succi-book} is $P=c_s^2\rho$, as it was shown in equation (\ref{navier-stokes:with:everything}), where $c_s$ is the scaling parameter which for the D2Q9 model equals $c_s=1/\sqrt{3}$.
-However, the inclusion of a density-dependent force term brings about a non-ideal contribution.
-Using the Taylor expansion, the Shan-Chen force (\ref{Shan-Chen:Shan-Chen:cont}) for the D2Q9 model can be expressed as follows:
-\begin{equation}
-\begin{aligned}
-\bm{F}=&G\psi(x,y)(w_1 \psi(x+c\Delta t,y)\bm{c_{1}}+w_2 \psi(x,y+c\Delta t)\bm{c_{2}}%
-+w_3 \psi(x-c\Delta t,y)\bm{c_{3}}+ \\
-&w_4 \psi(x,y-c\Delta t)\bm{c_{4}}+w_5 \psi(x+c\Delta t,y+c\Delta t)\bm{c_{5}}+ \\
-&w_6 \psi(x-c\Delta t,y+c\Delta t)\bm{c_{6}}+w_7 \psi(x-c\Delta t,y-c\Delta t)\bm{%
-c_{7}}+ \\
-&w_8 \psi(x+c\Delta t,y-c\Delta t)\bm{c_{8}})=G\psi(\frac{1}{3}\bm{\nabla}\psi+%
-\frac{1}{18}\bm{\nabla}\Delta\psi).
-\end{aligned}
-\end{equation}
-The following formulation is derived algebraically:
-\begin{equation}
-\label{shan-chen:force:taylor:expansion}
-\begin{aligned}
-F_{\alpha}&=G\psi\left(\frac{1}{3}\partial_{\alpha}\psi+\frac{1}{18}\partial_{\alpha}\Delta%
-\psi\right) =G\left(\frac{1}{6}\partial_{\alpha}\psi^2+\frac{1}{18}%
-\biggl(\partial_{\alpha}(\psi\Delta\psi)-\Delta\psi\partial_{\alpha}\psi\biggr)\right) \\
-&=G\left(\frac{1}{6}\partial_{\alpha}\psi^2+\frac{1}{18}%
-\biggl(\partial_{\alpha}(\psi\Delta\psi)+\frac{1}{2}\partial_{\alpha}|\bm{\nabla}\psi|^2
--\partial_{\beta}\partial_{\alpha}\psi\partial_{\beta}\psi\biggr)\right).
-\end{aligned}
-\end{equation}
-The force influence can be included to the momentum-flux tensor \cite{Sbragaglia-Succi}:
-\begin{equation*}
-\partial_{\beta}P_{\alpha\beta}=-F_{\alpha}+\partial_{\alpha}p=-F_{\alpha}+\partial_{\alpha}(c_s^2%
-\rho).
-\end{equation*}
-Thus, the flux tensor $P_{\alpha\beta}$ is modified as follows:
+The force modifies the momentum fluxes as:
 \begin{equation*}
@@
 |\bm{\nabla}\psi|^2 +\frac{G}{18}\psi\Delta\psi\biggr)\delta_{\alpha\beta}-\frac{G}{%
-18}\partial_{\alpha}\psi\partial_{\beta}\psi.
+18}\partial_{\alpha}\psi\partial_{\beta}\psi,
 \end{equation*}
-By analogy with classical mechanics, the potential of the force can be introduced as:
+where the equation of states is changed from original $P=c_s^2 \rho$ \cite{Succi-book} to the van-der-Waals type equation of state:
 \begin{equation} 
-\label{force:potential}
-U=\frac{G}{6}\psi^2+\frac{G}{36}|\bm{\nabla}\psi|^2+\frac{G}{18}%
-\psi\Delta\psi.
+P=c_s^2 \rho +\frac{G}{6} \psi(\rho)^2.
 \end{equation}
 
-Since the gradient terms in equation (\ref{force:potential}) are small compared with the leading terms (the characteristic length of the interface is longer than the lattice spacing, as in
-all diffuse-interface methods), the equation (\ref{force:potential}) can be approximated as
+{\color{red} Michal, you can add here binary Shan-Chen model. You have more information on it than me.}
+
+\subsection{Binary liquid multiphase model}
+In this section we'd like to describe binary liquid model. Binary model is based on the free energy model \cite{swift,landau}:
 \begin{equation}
-p=\rho c_s^2+\frac{G}{6}\psi^2.
-\end{equation}
-
-\subsection{Binary liquid model}
-\section{Binary liquid multiphase model}
-In this section we'd like to describe binary liquid model. Binary model is based on free energy model. The free energy for binary liquids can be written as \cite{swift,landau}:
-\begin{equation}
-F=\int{J \mathrm{d}V}=\int{\left(c_s^2\rho\ln\rho+\frac{A}{2}\phi^2+\frac{B}{4}\phi^4 + \frac{k}{2}(\nabla \phi)^2 \right)\mathrm{d}V} \quad,
+\mathfrak{F}=\int{J \mathrm{d}V}=\int{\left(c_s^2\rho\ln\rho+\frac{A}{2}\phi^2+\frac{B}{4}\phi^4 + \frac{k}{2}(\nabla \phi)^2 \right)\mathrm{d}V},
 \end{equation}
 where coefficients $A$ and $B$ are responsible for the phase separation, coefficient $k$ is responsible for the surface tension between phases and $\phi$ is the order parameter.
-Notice that if $A<0$ then the phase separation occurs, and $\phi$ will take values as $-1$ and $1$ for equilibrium phases.
-
-By minimising free energy functional with constraints for constant number of particles, one can obtain the chemical potential:
+Different phases are corresponding to extreme values of $\phi$ as $-1$ and $1$.
+The lattice Boltzmann model for the binary free-energy model is defined through two distribution sets:
 \begin{equation}
-\mu=A\phi+B\phi^3 - k \Delta \phi
+\begin{aligned}
+&f_i^{eq}=w_i \rho \biggl(1+\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\biggr), i=0\div8\\
+&g_i^{eq}=w_i(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi^m \frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}), i\neq0 \\
+&g_0^{eq}=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
+\end{aligned}
 \end{equation}
-From free energy model we can construct the pressure tensor:
+where $\Gamma$ is the mobility parameter, chemical potential $\mu=-A\phi+A\phi^3-k\Delta\phi$, $k$ is related to the surface tension. The macroscopic parameters are calculated as:
 \begin{equation}
-P_{\alpha \beta}=\bigg(p_0 - k\phi\Delta\phi-k/2 |\nabla \phi|^2\bigg)\delta_{\alpha\beta}+k\partial_{\alpha}\phi\partial_{\beta}\phi \quad,
+\begin{aligned}
+\rho=\sum_i{f_i}\\
+\rho u_{\alpha}=\sum_i{f_i c_{i\alpha}}\\
+\phi=\sum_i{g_i}
+\end{aligned}
 \end{equation}
-where
-\begin{equation}
- p_0=\rho c_s^2+\frac{A}{2}\phi^2+\frac{3}{4}B \phi^2
-\label{binary:bulk:pressure}
-\end{equation}
-
-Let us take parameter $A$ as $-B$, then after substitution we will obtain the following bulk pressure:
-\begin{equation}
- p_0=\rho c_s^2-\frac{B}{2}\phi^2+\frac{3}{4}B \phi^2
-\end{equation}
-
-The whole system of equations which describe binary model is the following:
+The equilibrium distribution functions lead to the macroscopic equations {\color{red} Give citation here}:
 \begin{equation}
@@
 &-\partial_{\alpha}P_{\alpha \beta} + \nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha} + \frac{1}{3}\partial_{\gamma} u_{\gamma} \delta_{\alpha\beta}\right)\\
-&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=D\Delta \mu
+&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=D\Delta \mu,
 \end{aligned}
@@
 \end{equation}
+with the bulk pressure $p_0=\rho c_s^2-\frac{B}{2}\phi^2+\frac{3}{4}B \phi^2$.
 
+{\color{red} Michal, need your help here. Which parameters have you used for simulation. Please check above.}
 We take equilibrium distribution function for the binary-liquid gas model. In general, it can be presented in the following form:
-\begin{equation}
+\subsection{Lattice Boltzmann implementation}
+For all distribution functions the collision and streaming step are performed:
+\begin{equation*}
 \begin{aligned}
-&g_i^{eq}=w_i(\Gamma \mu + \frac{\phi^m c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi^m \frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}), i\neq0 \\
-&g_0^{eq}=\phi^m-\sum_{i\neq0}{g_i^{eq}}\quad,
+f_i^{\star}=(1-\omega)f_i+f_i^{eq}\\
+f_i(\bm{x}+\bm{c_i})=f_i^{\star}. 
 \end{aligned}
-\end{equation}
-where $\mu$ is the chemical potential defined above, and $\Gamma$ is the mobility parameter.
+\end{equation*}
 

### 2010-04-04 paper.tex
Add more keywords, heavily update the abstract.

@@
 \begin{abstract}
-Despite the number and successful implementations of number of lattice Boltzmann systems on Graphics Processing Units (GPUs), the implementation of more complicated problems as multiphase models or Nearest-Neighbours (NN) interaction, which includes all the boundary conditions, on GPUs is still not well developed.
-The purpose of the paper is twofold: to develop the efficient GPUs implementation algorithms for the binary liquids flows, which are subclass of the NN interactions, and to illustrate the usefullness in terms of performance of GPUs for multiphase models simulations.
+Graphics Processing Units (GPUs) have been succesfully used to simulate
+various systems using the Lattice Boltzmann (LB) method.  The simulations were
+usually either a straightforward port of a CPU code (without any
+regards to performance considerations), or optimized for single phase
+flows.  The implementation of more complicated problems such as
+multiphase models or models with Nearest-Neighbours (NN) interaction
+has never received any special attention.  The purpose of this paper
+is twofold: to illustrate the usefulness of GPUs for multiphase simulations,
+and to carefully develop an optimized and efficient algorithm for GPU
+simulations of fluid flows using LB models with NN terms.
 \end{abstract}
@@
 \ams{52B10, 65D18, 68U05, 68U07{\color{red} We need to find somewhere classifications.}}
-\keywords{Lattice Boltzmann method, GPU, Shan-Chen model, Free-energy model, Binary liquids}
+\keywords{Lattice Boltzmann method, GPU, Shan-Chen model, Free-energy model, binary liquids, multiphase flow, CUDA, OpenCL, GPGPU}
 

### 2010-04-04 paper.tex
Break lines in the 1st paragraph.
Style fixes in the 1st paragraph.

@@
 \label{sec:introduction}
-During the last two decades since the lattice Boltzmann method was introduced \cite{mcnamara}, the method became mature to compete with the well established computational fluid dynamics (CFD).
-Due to the kinetic nature of the method the method is applicable to simulate not only hydrodynamics but also thermal flows \cite{karlin-minimalmodels,yuan-thermal}, micro-flows \cite{ansumali-small-knudsen}, ferrofluids \cite{rosensweig,kuzmin-aniso}, multiphase flows \cite{rothman-color,Shan-chen:extended,swift}.
-In the presence of complex physical effects such as turbulence, multi-phase flows, flow of oil and gas through porous material, flow around surfaces with complex geometries, or blood flow with deformable particles, the solution of the basic Navier-Stokes equations becomes difficult or impractical.
-On the other hand, the micro-scale calculations for such systems based on the  Molecular Dynamics approach (MD) are computationally extremely demanding.
-The Lattice Boltzmann Method (LBM), which focuses on the meso-scale, may be the best choice for such complex flow problems, being able to capture macro- and micro-flows.
+During the last two decades since its introduction \cite{mcnamara}, the Lattice Boltzmann Method (LBM)
+became mature enough to compete with other well-established approaches to computational fluid dynamics (CFD).
+Due to its kinetic nature, LBM can be used to simulate not only simple
+hydrodynamics but also thermal flows \cite{karlin-minimalmodels,yuan-thermal},
+micro-flows \cite{ansumali-small-knudsen}, ferrofluids \cite{rosensweig,kuzmin-aniso}, and
+multiphase flows \cite{rothman-color,Shan-chen:extended,swift}. In the presence of complex
+physical effects such as turbulence, multi-phase flows, flow of oil and gas through porous
+material, flow around surfaces with complex geometries, or blood flow with deformable
+particles, the solution of the basic Navier-Stokes equations becomes difficult or
+impractical. On the other hand, micro-scale calculations for such systems based
+on the Molecular Dynamics approach are computationally extremely demanding.
+The Lattice Boltzmann Method (LBM), which focuses on the meso-scale, may be the best
+choice for such complex flow problems, being able to capture the salient details of both macro- and micro-flows.
 

### 2010-04-04 paper.tex
Break lines in 3 more paragraphs.
Rewrite parts of the introduction section to make it shorter.
Also add info about Succi's GPU paper.

@@
 on the Molecular Dynamics approach are computationally extremely demanding.
-The Lattice Boltzmann Method (LBM), which focuses on the meso-scale, may be the best
+The lattice-Boltzmann method, which focuses on the meso-scale, may be the best
 choice for such complex flow problems, being able to capture the salient details of both macro- and micro-flows.
 
-{\color{red} Michal, please add here more details.}
-The lattice Boltzmann method is the explicit numerical method opposing some restrictions on the time propagation step.
-Thus the physical time step corresponding to the numerical time step is relatively small.
-Performing large-time-scale simulations become unpractical due to the high computational demand.
-To overcome the problem of the high computational demand the physical problems are calculated with the help of parallel processors of a cluster.
-The recent years trend is to perform the simulations on Graphical Processing Units (GPUs) which can be viewed as small supercomputers due to the large number of independent threads performing a simulation.
-The attractive part of simulations on GPUs is the low price (in comparison with clusters) and large performance.
-Not all the problems are matched to be simulated on parallel machines.
-However, the lattice Boltzmann method is a perfect match for parallel structure machines as GPUs.
+LB is an explicit numerical method, which poses some restrictions on the size of the time step --
+it has to be kept relatively small in order to preserve the numerical stability of the scheme.  Long-time
+simulations in large domains can thus quickly become impractical due to the high computational demand.
+The usual solution to this problem is to use a cluster to run the simulations.  In the last few
+years, an interesting alternative in the form of Graphics Processing Units of commodity video
+cards has emerged.  The GPUs can be viewed as small supercomputers due to their high
+computational power (e.g. 1 TFLOPS for a Tesla C1060) and ability to execute multiple
+independent threads simultaneously.  Their low price makes them an interesting hardware
+platform for numerical calculations of many kinds, but it has to be kept in mind that not
+all problems are well suited for the specific architecture of the GPU.  The Lattice-Boltzmann
+method, due it its inherent parallelism and locality is however a perfect match for this
+new hardware environment.
 
-During the last years the method was applied for many different problems, such as, to name a few, the calculation of the drag coefficient for the two-dimensional flow through cylinders \cite{tolke-twod} and moving sphere in three-dimensional channel \cite{tolke-GPU}.
-To the authors' best knowledge only simple hydrodynamics problems GPUs implementation is thoroughly described in the literature.
-However, the nearest neighbours (NN) interaction which includes advanced boundary lattice Boltzmann methods and multiphase problems are still not covered.
-For many multiphase problems the local density gradient is required to properly simulate the physics.
-The implementation of the multiphase models are more challenging than simple hydrodynamics problems because of the additional information required from other nodes.
-The simple hydrodynamics lattice Boltzmann method is local and is perfectly parallelized.
-The NN problems requires the additional information from other nodes and can't be addressed as local any more.
-
-This work's aim is to present the efficient GPUs algorithms to implement the binary liquids models based on the free-energy approach and the Shan-Chen method.
-First, we thoroughly describe the GPUs algorithms.
-Then the Shan-Chen model and the free-energy binary liquid model lattice Boltzmann implementation is presented.
-The numerical results for the comparison benchmark between those two models is done in terms of performance and accuracy in the ``Numerical results`` section.
-We conclude the paper with the main findings.
+This was quickly noticed by the CFD community and in the last few years the method was
+implemented on NVIDIA GPUs and applied to many different problems, such as calculation of
+the drag coefficient for the two-dimensional flow through cylinders \cite{tolke-twod} and
+for a moving sphere in a three-dimensional channel \cite{tolke-GPU}.  In the majority of these
+problems only fully local, single-fluid LB models were used.  Only very recently some
+work on more complex models with NN interactions has been done {\color{red} cite Succi's paper},
+but using a code ported from a CPU version, and not optimized specifically for the GPU
+environment.
+%%
+% To the authors' best knowledge only simple
+% hydrodynamics problems GPUs implementation is thoroughly described in the literature.
+%%
+% This is actually not true anymore -- see the Succi's PRE paper about LBM on GPUs
+The aim of this paper is to present efficient GPU-optimized algorithms for LB models
+with NN interaction terms.  Such models are abundant and of enormous practical importance.
+They include the Shan-Chen model, free-energy binary fluid models and higher-order
+boundary conditions for fluid flows.  In the following sections, we first present
+the details of the Shan-Chen and free-energy LB models.  The reader
+is then briefly introduced to the main features of the CUDA environment
+that we use to program the GPUs, which is followed by a thorough description
+of the GPU algorithms and an analysis of their efficiency.  In section \ref{sec:bechmark}, the two
+binary fluid models are compared in terms of performance and accuracy.
+We conclude the paper with a summary of the main findings.
 
@@
 \section{Results}
+\label{sec:benchmark}
+
 \section{Conclusion}

### 2010-04-05 paper.tex
Break lines in the remainder of the text.

@@
 
-\section{LBM}
+\section{The Lattice-Boltzmann method}
 \label{sec:lbm:binary:liquids}
-This works concentrates on two popular LBM multiphase models.
-Originally introduced in 1993 \cite{Shan-chen:extended} the model attracted a multitude of researchers by its simplicity.
-The free-energy model \cite{swift} is based on the free-energy phase transition \cite{landau}.
-Both models werre successfully applied in simulation of two-phase blood flow \cite{halliday-multicomponent}, flow in microchannels \cite{pooley-contact}, thermal flow \cite{zhang-thermal}, and studies of break-up of the liquid droplet \cite{nourgaliev-breakup}.
-The section describes the two-dimensional implementation of the Shan-Chen and free-energy models with the D2Q9 model.
+This works concentrates on two popular LBM multiphase models. Originally introduced
+in 1993 \cite{Shan-chen:extended} the model attracted a multitude of researchers
+by its simplicity.  The free-energy model \cite{swift} is based on the free-energy
+phase transition \cite{landau}. Both models werre successfully applied in simulation
+of two-phase blood flow \cite{halliday-multicomponent}, flow in microchannels
+\cite{pooley-contact}, thermal flow \cite{zhang-thermal}, and studies of break-up
+of the liquid droplet \cite{nourgaliev-breakup}. The section describes the
+two-dimensional implementation of the Shan-Chen and free-energy models with
+the D2Q9 model.
 
 \subsection{Shan-Chen model}
-In the Shan-Chen model \cite{Shan-chen:extended} the force at a given node depends on all local neighbours characteristics:
+In the Shan-Chen model \cite{Shan-chen:extended} the force at a given node depends
+on all local neighbours characteristics:
 \begin{equation}  \label{Shan-Chen:Shan-Chen:cont}
@@
 \end{equation}
-where $\psi$ is the function of local characteristics of a node, and taken as $\psi(\rho)=1-\exp(-\rho)$.
-Coefficient $G$ is responsible for the attraction force between the molecules.
-Overall, the Shan-Chen force is the approximation of the force between molecules \cite{kwok,kwok-contact-angle}.
-The force modifies the momentum fluxes as:
+where $\psi$ is the function of local characteristics of a node, and taken
+as $\psi(\rho)=1-\exp(-\rho)$. Coefficient $G$ is responsible for the attraction
+force between the molecules. Overall, the Shan-Chen force is the approximation of
+the force between molecules \cite{kwok,kwok-contact-angle}. The force modifies
+the momentum fluxes as:
 \begin{equation*}
@@
 \end{equation*}
-where the equation of states is changed from original $P=c_s^2 \rho$ \cite{Succi-book} to the van-der-Waals type equation of state:
+where the equation of states is changed from original $P=c_s^2 \rho$ \cite{Succi-book}
+to the van-der-Waals type equation of state:
 \begin{equation}
@@
 \subsection{Binary liquid multiphase model}
-In this section we'd like to describe binary liquid model. Binary model is based on the free energy model \cite{swift,landau}:
+In this section we describe the binary liquid model. Binary model is based on
+the free energy model \cite{swift,landau}:
 \begin{equation}
@@
 \end{equation}
-where coefficients $A$ and $B$ are responsible for the phase separation, coefficient $k$ is responsible for the surface tension between phases and $\phi$ is the order parameter.
-Different phases are corresponding to extreme values of $\phi$ as $-1$ and $1$.
-The lattice Boltzmann model for the binary free-energy model is defined through two distribution sets:
+where coefficients $A$ and $B$ are responsible for the phase separation,
+coefficient $k$ is responsible for the surface tension between phases and
+$\phi$ is the order parameter. Different phases are corresponding to
+extreme values of $\phi$ as $-1$ and $1$.
+The lattice Boltzmann model for the binary free-energy model is defined
+through two distribution sets:
 \begin{equation}
@@
 \end{equation}
-where $\Gamma$ is the mobility parameter, chemical potential $\mu=-A\phi+A\phi^3-k\Delta\phi$, $k$ is related to the surface tension. The macroscopic parameters are calculated as:
+where $\Gamma$ is the mobility parameter, chemical potential
+$\mu=-A\phi+A\phi^3-k\Delta\phi$, $k$ is related to the surface
+tension. The macroscopic parameters are calculated as:
 \begin{equation}
@@
 \end{equation}
-The equilibrium distribution functions lead to the macroscopic equations {\color{red} Give citation here}:
+The equilibrium distribution functions lead to the macroscopic equations
+{\color{red} Give citation here}:
 \begin{equation}
@@
 We take equilibrium distribution function for the binary-liquid gas model. In general, it can be presented in the following form:
+
 \subsection{Lattice Boltzmann implementation}

### 2010-04-05 paper.tex
Start a section about the CUDA programming model.

@@
 \section{The Lattice-Boltzmann method}
+
+{\color{red} a general overview of basics of LBM needs to be added here (grid, collision, propagation, etc)}
+
 \label{sec:lbm:binary:liquids}
@@
 
-\section{GPU implementation}
+\section{The CUDA Programming Environment}
+
+CUDA is implemented in the video card hardware by organizing the GPU around the
+concept of a multiprocessor (MP).  Such a multiprocessor consists of several (8 for
+devices of compute capability lower than 2.0) scalar processors (SPs), each of which is
+capable of executing a thread in a SIMT (Single Instruction, Multiple Threads) manner.
+Each MP also has a limited amount of specialized on-chip memory: a set of 32-bit registers,
+a shared memory block, a constant cache and a texture cache.  The registers are local
+to the scalar processor, but the other types of memory are shared between all SPs
+in a MP, which allows data sharing between threads.
+
+{\color{red} TODO: verify the figures in this paragraph}
+Perhaps the most salient feature of the CUDA architecture is the memory hierarchy
+with 1-2 orders of magnitude differences between access times at each successive levels.
+
+The slowest kind of memory is the host memory accessible only to the CPU.  This memory,
+while potentially very large in size, cannot be accessed directly by the GPU, so in order
+for data stored there to be useful during a GPU simulation, it has to be copied to the
+global device memory.  Such a data transfer goes over the PCI-e bus, which has a maximum
+throughput of about 6~GB/s.
+
+Next in line is the global device memory of the GPU, which can currently span several
+gigabytes and which has a bandwidth of about 100~GB/s.  Global memory accesses are however
+high-latency operations, taking several hundred clock cycles of the GPU to complete {\color{red} Cite NVIDIA documentation here}
+
+The fastest kind of memory currently available on GPUs is the shared memory block residing
+on MPs.  It is currently limited in size to just 16~kB, but has a bandwidth of ca~1.3~TB/s
+and usually a latency no higher than that of a~SP register access.
+
+The above description readily suggests an optimization strategy which we will generally
+follow in the next section and which can be summarized as: move as much data as possible
+to the fastest kind of memory available and keep it there as long as possible, while minimizing
+accesses to the slower kinds of memory.
+
+\section{GPU implementation of LBM}
+
+We will now outline an efficient and general GPU LBM algorithm taking into consideration
+the facts from the previous section.
+
 \section{Results}

### 2010-04-07 misc/efficiency.py
Add some internal notes about performance.

@@ -1,0 +1,25 @@
+#!/usr/bin/python
+
+def sc2d(n):
+    print 'default ', 37.0 * n**2 / (9 * n**2)
+    print 'shared X', 31.0 * n**2 / (9 * n**2)
+    print 'two pass', (10.0 * (n-1)**2 + (4*n-4) + 19 * (n-1)**2) / (9.0*(n-1)*(n-1))
+    print 'one pass', (9.0 * n**2 + 9 * (n-1)**2) / (9 * (n-1)**2)
+
+def sc3d(n):
+    print 'default ', 77.0 * n**2 / (19 * n**2)
+    print 'two pass', (20.0*2 * (n-1)**2 + (4*n-4) + 19*(n-1)**2) / (19 * (n-1)**2)
+    print 'one pass', (19.0 * n**2 + 19 * (n-1)**2) / (19 * (n-1)**2)
+    print '2 one pass', (19.0*2 * n**2 + 19*2*(n-1)**2) / (19 * 2 * (n-1)**2)
+
+print '* sc2d'
+for x in [16, 24, 32, 40, 48, 56]:
+    print '**', x
+    sc2d(x)
+    print
+
+print '* sc3d'
+for x in [16, 24, 32]:
+    print '**', x
+    sc3d(x)
+    print

### 2010-04-08 codes/ShanChenSucci/cpu.cpp
Added stencils paper. Added the Shan-Chen GPU with VdW analysis. Added Succi's model, but haven't checked against the Maxwellian area rule.

@@ -1,0 +1,190 @@
+/*
+ *  cpu.cpp
+ *  OpenCLTest
+ *
+ *  Created by Alexandr Kuzmin on 09-09-26.
+ *  Copyright 2009 __MyCompanyName__. All rights reserved.
+ *
+ */
+
+/*
+ *  main.cpp
+ *  OpenCLTest
+ *
+ *  Created by Alexandr Kuzmin on 09-09-24.
+ *  Copyright 2009 __MyCompanyName__. All rights reserved.
+ *
+ */
+
+//#include "main.h"
+////////////////////////////////////////////////////////////////////////////////
+
+#include <iostream>
+#include <cmath>
+#include <time.h>
+#include <fstream>
+#include <sstream>
+#include <string>
+#include <cstdlib>
+#include <algorithm>
+////////////////////////////////////////////////////////////////////////////////
+
+const int nx=256;
+const int ny=256;
+
+void writeFile(std::string name, float* f, int n)
+{
+	std::ofstream fout(name.c_str());
+	for(int counterX=0;counterX<nx;counterX++)
+	{	
+		for(int counterY=0;counterY<ny;counterY++)
+			fout<<f[counterX*ny+counterY]<<" ";
+		fout<<"\n";
+	}		
+	fout<<std::endl;
+}
+int main(int argc, char** argv)
+{
+    int err;                            // error code returned from api calls
+	const int npop = 9;
+	
+	float rho[nx*ny];
+	float u1[nx*ny];
+	float u2[nx*ny];
+	float f_mem[nx*ny*npop];
+	float f2_mem[nx*ny*npop];
+	float feq[npop];
+	
+	float gradlaplace[5][5]={{-1.0/72.0,-4.0/72.0,0.0,4.0/72.0,1.0/72.0},
+				 {-8.0/72.0,4.0/72.0,0.0,-4.0/72.0,8.0/72.0},
[…]

### 2010-04-08 codes/ShanChenVdW/main.cpp

@@ -1,0 +1,577 @@
+/*
+ *  main.cpp
+ *  Shan-Chen GPU code
+ *
+ *  Created by Alexandr Kuzmin on 09-09-24.
+ *  Copyright 2009 __MyCompanyName__. All rights reserved.
+ *
+ */
+
+////////////////////////////////////////////////////////////////////////////////
+
+#include <sys/types.h>
+#include <sys/stat.h>
+#include <time.h>
+#include <CL/cl.h>
+#include <iostream>
+#include <cmath>
+#include <fstream>
+#include <sstream>
+#include <string>
+#include <cstdlib>
+
+#define BLOCK_SIZE 64
+////////////////////////////////////////////////////////////////////////////////
+////////////////////////////////////////////////////////////////////////////////
+static char *
+load_program_source(const char *filename)
+{
+    struct stat statbuf;
+    FILE        *fh;
+    char        *source;
+	
+    fh=fopen(filename, "r");
+	if (fh == 0)
+	{
+		return 0;
+	}
+    stat(filename, &statbuf);
+    source = (char *) malloc(statbuf.st_size + 1);
+    fread(source, statbuf.st_size, 1, fh);
+    source[statbuf.st_size] = '\0';
+	
+    return source;
+}
+
+void writeFile(std::string name, float* f, int nx,int ny)
+{
+	std::ofstream fout(name.c_str());
+	for(int counterX=0;counterX<nx;counterX++)
+	{
+		for(int counterY=0;counterY<ny;counterY++)
+			fout<<f[counterX*ny+counterY]<<" ";
+		fout<<"\n";
+	}
+	fout<<std::endl;
+}
+
+int main(int argc, char** argv)
+{
+    int err;                            // error code returned from api calls
+	int nx = 256;
+	int ny=256;
+	double rhol=4.2;
+	double rhog=2.8;
+	double t=4.0/7.0*0.99;
+	double g=-1.0;
+	double tau=1.0;
+	int radius=40;
+	
+	float host_fr0[nx*ny];
+	float host_fe0[nx*ny];
+	float host_fn0[nx*ny];
+	float host_fw0[nx*ny];
[…]

### 2010-04-08 codes/ShanChenSucci/cpu_simple.cpp
Added the Mathematica file with calculations of bulk densities. Succi's model results are still to come.

@@ -1,0 +1,240 @@
+/*
+ *  cpu.cpp
+ *  OpenCLTest
+ *
+ *  Created by Alexandr Kuzmin on 09-09-26.
+ *  Copyright 2009 __MyCompanyName__. All rights reserved.
+ *
+ */
+
+/*
+ *  main.cpp
+ *  OpenCLTest
+ *
+ *  Created by Alexandr Kuzmin on 09-09-24.
+ *  Copyright 2009 __MyCompanyName__. All rights reserved.
+ *
+ */
+
+//#include "main.h"
+////////////////////////////////////////////////////////////////////////////////
+
+#include <iostream>
+#include <cmath>
+#include <time.h>
+#include <fstream>
+#include <sstream>
+#include <string>
+#include <cstdlib>
+#include <algorithm>
+////////////////////////////////////////////////////////////////////////////////
+
+const int nx=256;
+const int ny=256;
+
+void writeFile(std::string name, float* f, int n)
+{
+	std::ofstream fout(name.c_str());
+	for(int counterX=0;counterX<nx;counterX++)
+	{	
+		for(int counterY=0;counterY<ny;counterY++)
+			fout<<f[counterX*ny+counterY]<<" ";
+		fout<<"\n";
+	}		
+	fout<<std::endl;
+}
+int main(int argc, char** argv)
+{
+    int err;                            // error code returned from api calls
+	const int npop = 9;
+	
+	float rho[nx*ny];
+	float u1[nx*ny];
+	float u2[nx*ny];
+	float f_mem[nx*ny*npop];
+	float f2_mem[nx*ny*npop];
+	float feq[npop];
+	
+	float gradlaplace[25]={-1.0/72.0,-4.0/72.0,0.0,4.0/72.0,1.0/72.0,
+				 -8.0/72.0,4.0/72.0,0.0,-4.0/72.0,8.0/72.0,
[…]

### 2010-04-08 codes/ShanChenSucci/cpu.cpp
Mathematica files should be OK right now.

@@ -55,17 +55,33 @@ int main(int argc, char** argv)
 	float f2_mem[nx*ny*npop];
 	float feq[npop];
 	
-	float gradlaplace[5][5]={{-1.0/72.0,-4.0/72.0,0.0,4.0/72.0,1.0/72.0},
-				 {-8.0/72.0,4.0/72.0,0.0,-4.0/72.0,8.0/72.0},
-				 {-18.0/72.0,72.0/72.0,0.0,-72.0/72.0,18.0/72.0},
-				 {-8.0/72.0,4.0/72.0,0.0,-4.0/72.0,8.0/72.0},
-				 {-1.0/72.0,-4.0/72.0,0.0,4.0/72.0,1.0/72.0}};
-	 
-	int stencil[25][5]={{-2,2},{-1,2},{0,2},{1,2},{2,2},
+	float gradlaplace[25]={-1.0/72.0,-4.0/72.0,0.0,4.0/72.0,1.0/72.0,
+				 -8.0/72.0,4.0/72.0,0.0,-4.0/72.0,8.0/72.0,
+				 -18.0/72.0,72.0/72.0,0.0,-72.0/72.0,18.0/72.0,
+				 -8.0/72.0,4.0/72.0,0.0,-4.0/72.0,8.0/72.0,
+				 -1.0/72.0,-4.0/72.0,0.0,4.0/72.0,1.0/72.0};
+
+	float gradstencilx[9]={0.0,4.0/12.0,0.0,-4.0/12.0,0.0,
+			          1.0/12.0,-1.0/12.0,-1.0/12.0,1.0/12.0};
+
+	float gradstencily[9]={0.0,0.0,4.0/12.0,0.0,-4.0/12.0,
+			          1.0/12.0,1.0/12.0,-1.0/12.0,-1.0/12.0};
+
+
+	float laplacestencil[9]={-20.0/6.0,4.0/6.0,4.0/6.0,4.0/6.0,4.0/6.0,
+					   1.0/6.0,1.0/6.0,1.0/6.0,1.0/6.0};
+
+	int stencil[25][2]={{-2,2},{-1,2},{0,2},{1,2},{2,2},
 			    {-2,1},{-1,1},{0,1},{1,1},{2,1},
 			    {-2,0},{-1,0},{0,0},{1,0},{2,0},
 			    {-2,-1},{-1,-1},{0,-1},{1,-1},{2,-1},
 			    {-2,-2},{-1,-2},{0,-2},{1,-2},{2,-2}};
+
[…]

### 2010-04-08 codes/ShanChenSucci/cpu_simple.cpp

@@ -55,32 +55,6 @@ int main(int argc, char** argv)
 	float f2_mem[nx*ny*npop];
 	float feq[npop];
 	
-	float gradlaplace[25]={-1.0/72.0,-4.0/72.0,0.0,4.0/72.0,1.0/72.0,
-				 -8.0/72.0,4.0/72.0,0.0,-4.0/72.0,8.0/72.0,
-				 -18.0/72.0,72.0/72.0,0.0,-72.0/72.0,18.0/72.0,
-				 -8.0/72.0,4.0/72.0,0.0,-4.0/72.0,8.0/72.0,
-				 -1.0/72.0,-4.0/72.0,0.0,4.0/72.0,1.0/72.0};
-
-	float gradstencilx[9]={0.0,4.0/12.0,0.0,-4.0/12.0,0.0,
-			          1.0/12.0,-1.0/12.0,-1.0/12.0,1.0/12.0};
-
-	float gradstencily[9]={0.0,0.0,4.0/12.0,0.0,-4.0/12.0,
-			          1.0/12.0,1.0/12.0,-1.0/12.0,-1.0/12.0};
-
-
-	float laplacestencil[9]={-20.0/6.0,4.0/6.0,4.0/6.0,4.0/6.0,4.0/6.0,
-					   1.0/6.0,1.0/6.0,1.0/6.0,1.0/6.0};
-
-	int stencil[25][2]={{-2,2},{-1,2},{0,2},{1,2},{2,2},
-			    {-2,1},{-1,1},{0,1},{1,1},{2,1},
-			    {-2,0},{-1,0},{0,0},{1,0},{2,0},
-			    {-2,-1},{-1,-1},{0,-1},{1,-1},{2,-1},
-			    {-2,-2},{-1,-2},{0,-2},{1,-2},{2,-2}};
-
-	int smallstencil[9][2]={{-1,1},{0,1},{1,1},
-			        {-1,0},{0,0},{1,0},
-			        {-1,-1},{0,-1},{1,-1}};
-
 
 	float* f=f_mem;
 	float* f2=f2_mem;
@@ -167,28 +141,9 @@ int main(int argc, char** argv)
 				fy+=weights[k]*cy[k]*(1.0-exp(-rho[nx*iY2+iX2]));
 			}
 
-			for(int k=0;k<9;k++)
-			{
-				float lapl[9];
[…]

### 2010-04-08 codes/ShanChenSucci/Analysis/analysis.m
Added the Succi's model comparison with the usual model. There are no any seen differencies.

@@ -1,0 +1,14 @@
+clear all;
+clc;
+
+dense_succi=load('height09800_succi.dat');
+dense_normal=load('height09800.dat');
+
+%% Plotting data
+plot(1:256,dense_succi(:,128),1:256,dense_normal(:,128))
+
+%% Density data
+mean(dense_succi(180:220,128))
+mean(dense_normal(180:220,128))
+mean(dense_succi(50:100,128))
+mean(dense_normal(50:100,128))

### 2010-04-19 codes/JuliaShanChen/analysis.py
Added Julia's model with Shan-Chen potential. Verified coefficients with Julia's model. Can't stabilize the density plan simulation. Probably need to smooth the interface.

@@ -1,0 +1,9 @@
+import pylab
+import numpy
+import os
+
+print os.getcwd()
+profile=numpy.loadtxt("height00220.dat")
+
+pylab.plot(profile[128, :])
+pylab.show();

### 2010-04-19 codes/JuliaShanChen/cpu.cpp

@@ -1,0 +1,207 @@
+/*
+ *  cpu.cpp
+ *  OpenCLTest
+ *
+ *  Created by Alexandr Kuzmin on 09-09-26.
+ *  Copyright 2009 __MyCompanyName__. All rights reserved.
+ *
+ */
+
+/*
+ *  main.cpp
+ *  OpenCLTest
+ *
+ *  Created by Alexandr Kuzmin on 09-09-24.
+ *  Copyright 2009 __MyCompanyName__. All rights reserved.
+ *
+ */
+
+//#include "main.h"
+////////////////////////////////////////////////////////////////////////////////
+
+//#include <OpenCL/opencl.h>
+#include <iostream>
+#include <cmath>
+#include <time.h>
+#include <fstream>
+#include <sstream>
+#include <string>
+#include <cstdlib>
+#include <algorithm>
+////////////////////////////////////////////////////////////////////////////////
+
+const int nx=256;
+const int ny=256;
+
+void writeFile(std::string name, float* f, int n)
+{
+	std::ofstream fout(name.c_str());
+	for(int counterX=0;counterX<nx;counterX++)
+	{	
+		for(int counterY=0;counterY<ny;counterY++)
+			fout<<f[counterX*ny+counterY]<<" ";
+		fout<<"\n";
+	}		
+	fout<<std::endl;
+}
+int main(int argc, char** argv)
+{
+    int err;                            // error code returned from api calls
+	const int npop = 9;
+	
+	float lapl[nx*ny];
+	float gradx[nx*ny];
+	float grady[nx*ny];
+	float rho[nx*ny];
+	float u1[nx*ny];
+	float u2[nx*ny];
+	float f_mem[nx*ny*npop];
+	float f2_mem[nx*ny*npop];
+	float feq[npop];
+	
+	float* f=f_mem;
+	float* f2=f2_mem;
+	
[…]

### 2010-04-20 paper.tex
Added the one component description with Julia's model. Still cannot make the model working with the Shan-Chen potential even with smoothing.

@@
-\documentclass[mathpazo,sort,numbers]{cicp}
+\documentclass[mathpazo]{cicp}
 \usepackage{bm}
@@
 that we use to program the GPUs, which is followed by a thorough description
-of the GPU algorithms and an analysis of their efficiency.  In section \ref{sec:bechmark}, the two
+of the GPU algorithms and an analysis of their efficiency.  In section \ref{sec:benchmark}, the two
 binary fluid models are compared in terms of performance and accuracy.
@@
 the D2Q9 model.
+
+The lattice Boltzmann implementation in a standard view:
+\begin{equation}
+\label{standard:implementation}
+\begin{aligned}
+&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t)\\
+&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t). 
+\end{aligned}
+\end{equation}
 
@@
 We take equilibrium distribution function for the binary-liquid gas model. In general, it can be presented in the following form:
+\subsection{Free-energy model}
+The free-energy or pressure model \cite{swift} allows the incorporation of the any equation of state to the lattice Boltzmann equation.
+The lattice Boltzmann scheme is implemented in a standard form (\ref{standard:implementation}), but the equilibrium function is defined as:
+\begin{equation}
+\begin{aligned}
+f_i^{eq}=&w_i \biggl(p_0(\rho)-3 k\rho\Delta\rho+\rho \frac{c_{i\alpha}u_{\alpha}}{3}+\rho \frac{9 Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2}\biggr)\\
+&+k (w_{ixx} \partial_x \rho \partial_x \rho + w_{ixy} \partial_x \rho \partial_y \rho + w_{iyy} \partial_y \rho \partial_y \rho),
+\end{aligned}
+\end{equation}
+where $w_{ixx}=\{\dfrac{1}{3},\dfrac{1}{3},-\dfrac{1}{6},-\dfrac{1}{6},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24}\}$, $w_{iyy}=\{\dfrac{1}{3},\dfrac{1}{3},-\dfrac{1}{6},-\dfrac{1}{6},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24}\}$ and $w_{ixy}=\{0,0,0,0,\dfrac{1}{4},\dfrac{1}{4},-\dfrac{1}{4},-\dfrac{1}{4}\}$.
+The laplacian and the gradient are taken by the numerical stencils:
+\begin{equation}
+\Delta=\frac{1}{6}
+\begin{pmatrix}
+1&4&1\\
+4&-20&4\\
+1&4&1\\
+\end{pmatrix}\text{ and }
+\partial_x=\frac{1}{12}
+\begin{pmatrix}
+-1&0&1\\
+-4&0&4\\
+-1&0&1\\
+\end{pmatrix}.
+\end{equation}
+
+We implement in terms of convergence,accuracy and performance this model with the Shan-Chen potential, mainly $p_0=\dfrac{\rho}{3}+G\dfrac{\psi^2(\rho)}{6}$.
 
@@
 \begin{aligned}
-f_i^{\star}=(1-\omega)f_i+f_i^{eq}\\
-f_i(\bm{x}+\bm{c_i})=f_i^{\star}. 
+&f_i^{\star}=(1-\omega)f_i+f_i^{eq}\\
+&f_i(\bm{x}+\bm{c_i})=f_i^{\star}. 
 \end{aligned}
@@
 throughput of about 6~GB/s.
+% max 3.2 GB/s one way, measured on a 4xGTX285 system
 
@@
 high-latency operations, taking several hundred clock cycles of the GPU to complete {\color{red} Cite NVIDIA documentation here}
+% 131.9 GB/s for GTX 280, see NVIDIA CUDA Best Practices Guide, v3.0
 

### 2010-04-20 codes/JuliaShanChen/analysis.py

@@ -3,7 +3,7 @@ import numpy
 import os
 
 print os.getcwd()
-profile=numpy.loadtxt("height00220.dat")
+profile=numpy.loadtxt("height00000.dat")
 
-pylab.plot(profile[128, :])
+pylab.plot(profile[:, 128])
 pylab.show();

### 2010-04-20 codes/JuliaShanChen/cpu.cpp

@@ -63,12 +63,35 @@ int main(int argc, char** argv)
 	float* f2=f2_mem;
 	
 	float g=-5.0, tau=1.0;
-	float surf=0.05; //-1.0/(3.0*g);
+	float surf=0.01; //-1.0/(3.0*g);
 	int radius=40;
+	float width=5.0;
+	float sigma=1.0;
 	float rhol=1.93;
 	float rhog=0.16;
 	int i;
 	float dense,v1,v2;
+	
+	
+	
+	for (i=0; i<nx*ny; i++)
+	{
+		if (i/nx<=double(ny)/2.0)
+			rho[i]=rhol;
+		else 
+			rho[i]=rhog;
+	}
+	
+	for (int j=0;j<int(width);j++)
+	{
+		for(i=0;i<nx;i++)
+		{
+			rho[(ny/2+j)*nx+i]=rhog+(rhol-rhog)*exp(-j*j/(2.0*sigma*sigma));
+			rho[j*nx+i]=rhog+(rhol-rhog)*exp(-(j-width+1)*(j-width+1)/(2.0*sigma*sigma));
+		}
+	}
+	
+	
 	for (i=0; i<nx*ny; i++)
 	{	
 	/*	if ((i/nx-double(ny)/2.0)*(i/nx-double(ny)/2.0)+(i%nx-double(nx)/2.0)*(i%nx-double(nx)/2.0)<=radius*radius)
@@ -78,10 +101,6 @@ int main(int argc, char** argv)
 		else 
 			rho[i]=rhog;
 	*/	
-		if (i/nx>=double(ny)/2.0)
-			rho[i]=rhol;
-		else
-			rho[i]=rhog;
 
 		
 		dense=rho[i];

### 2010-04-21 paper.tex
Modified lattice Boltzmann part of the paper. Need more work with the description of the binary liquid models.

@@
 and to carefully develop an optimized and efficient algorithm for GPU
-simulations of fluid flows using LB models with NN terms.
+fluid flows simulations using LB models with NN terms.
 \end{abstract}
@@
 \label{sec:lbm:binary:liquids}
-This works concentrates on two popular LBM multiphase models. Originally introduced
+This work concentrates on two popular LBM multiphase models. Originally introduced
 in 1993 \cite{Shan-chen:extended} the model attracted a multitude of researchers
 by its simplicity.  The free-energy model \cite{swift} is based on the free-energy
-phase transition \cite{landau}. Both models werre successfully applied in simulation
-of two-phase blood flow \cite{halliday-multicomponent}, flow in microchannels
+phase transition model \cite{landau}. Both models were successfully applied in simulation
+of two-component blood flow \cite{halliday-multicomponent}, flow in microchannels
 \cite{pooley-contact}, thermal flow \cite{zhang-thermal}, and studies of break-up
@@
 
-The lattice Boltzmann implementation in a standard view:
+The lattice Boltzmann equation operates on the rectangular grid representing the physical domain.
+It utilizes the probability distribution functions containing the information about the macroscopic variables, such as density, momentum.
+LBE consists from two parts: local collision and propagation operated from one node to another in the certain direction specified by the velocity set.
+The lattice Boltzmann implementation in a standard view representing the collision and propagation steps:
 \begin{equation}
@@
 \begin{aligned}
-&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t)\\
-&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t). 
+&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t) + F_i,&&\text{ collision step}\\
+&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{ propagation step}, 
 \end{aligned}
 \end{equation}
+where $F_i$ is the external force population. The equilibrium function population is usually given as \cite{}:
+\begin{equation}
+f_i^{eq}=w_i \rho \biggl(1+\frac{c_{i\alpha}u_{\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\biggr),
+\end{equation}
+where for D2Q9 $w_i=\{\dfrac{4}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{36},\dfrac{1}{36},\dfrac{1}{36},\dfrac{1}{36}\}$ and $Q_{i\alpha\beta}=$ with the sound speed parameter $c_s^2=\dfrac{1}{3}$.
+The macroscopic variables are restored on every time step through the probability function population:
+\begin{equation*}
+\begin{aligned}
+&\rho=\sum_i{f_i}
+&\rho u_{\alpha}=\sum_i{f_i c_{i\alpha}}
+\end{aligned}
+\end{equation*}
+The multiphase models can be simulated either through the force population and/or special equilibrium function.
+The Shan-Chen model utilizes the force population, but the free-energy model is simulated through the special equilibrium model.
+Both models are described in the next sections.
 
@@
 \end{equation}
-
-{\color{red} Michal, you can add here binary Shan-Chen model. You have more information on it than me.}
-
-\subsection{Binary liquid multiphase model}
-In this section we describe the binary liquid model. Binary model is based on
-the free energy model \cite{swift,landau}:
+The lattice Boltzmann simulation for the Shan-Chen model is implemented through the standard implementation (\ref{standard:implementation}).
+In the presence of the external force the macroscopic velocity is changed as $\rho u_{\alpha}=\sum_i{f_i c_{i\alpha}}+\dfrac{F_{\alpha}}{2}$.
+The force population is defined as \cite{guo}:
 \begin{equation}
-\mathfrak{F}=\int{J \mathrm{d}V}=\int{\left(c_s^2\rho\ln\rho+\frac{A}{2}\phi^2+\frac{B}{4}\phi^4 + \frac{k}{2}(\nabla \phi)^2 \right)\mathrm{d}V},
+F_i=w_i\biggl(1-\frac{\omega}{2}\biggr)\biggl(3 (c_{i\alpha}-u_{\alpha}) + 9 (c_{i\beta}u_{\beta})c_{i\alpha}\biggr)F_{\alpha}.
 \end{equation}
-where coefficients $A$ and $B$ are responsible for the phase separation,
-coefficient $k$ is responsible for the surface tension between phases and
-$\phi$ is the order parameter. Different phases are corresponding to
-extreme values of $\phi$ as $-1$ and $1$.
-The lattice Boltzmann model for the binary free-energy model is defined
-through two distribution sets:
-\begin{equation}
-\begin{aligned}
-&f_i^{eq}=w_i \rho \biggl(1+\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\biggr), i=0\div8\\
-&g_i^{eq}=w_i(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi^m \frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}), i\neq0 \\
-&g_0^{eq}=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
-\end{aligned}
-\end{equation}
-where $\Gamma$ is the mobility parameter, chemical potential
-$\mu=-A\phi+A\phi^3-k\Delta\phi$, $k$ is related to the surface
-tension. The macroscopic parameters are calculated as:
-\begin{equation}
-\begin{aligned}
-\rho=\sum_i{f_i}\\
-\rho u_{\alpha}=\sum_i{f_i c_{i\alpha}}\\
-\phi=\sum_i{g_i}
-\end{aligned}
-\end{equation}
-The equilibrium distribution functions lead to the macroscopic equations
-{\color{red} Give citation here}:
-\begin{equation}
-\begin{aligned}
-&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
-&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=\\
-&-\partial_{\alpha}P_{\alpha \beta} + \nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha} + \frac{1}{3}\partial_{\gamma} u_{\gamma} \delta_{\alpha\beta}\right)\\
-&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=D\Delta \mu,
-\end{aligned}
-\label{binary:fluid:system}
-\end{equation}
-with the bulk pressure $p_0=\rho c_s^2-\frac{B}{2}\phi^2+\frac{3}{4}B \phi^2$.
-
-{\color{red} Michal, need your help here. Which parameters have you used for simulation. Please check above.}
-We take equilibrium distribution function for the binary-liquid gas model. In general, it can be presented in the following form:
+The Shan-Chen model is one-component multiphase model. The multicomponent modification is presented in the Appendix {\color{red} put appendix reference here}.
 \subsection{Free-energy model}
 The free-energy or pressure model \cite{swift} allows the incorporation of the any equation of state to the lattice Boltzmann equation.
+In comparison with the Shan-Chen model, the free-energy model is simulated through the special choice of the equilibrium distribution function.
 The lattice Boltzmann scheme is implemented in a standard form (\ref{standard:implementation}), but the equilibrium function is defined as:
@@
 We implement in terms of convergence,accuracy and performance this model with the Shan-Chen potential, mainly $p_0=\dfrac{\rho}{3}+G\dfrac{\psi^2(\rho)}{6}$.
-
-\subsection{Lattice Boltzmann implementation}
-For all distribution functions the collision and streaming step are performed:
-\begin{equation*}
-\begin{aligned}
-&f_i^{\star}=(1-\omega)f_i+f_i^{eq}\\
-&f_i(\bm{x}+\bm{c_i})=f_i^{\star}. 
-\end{aligned}
-\end{equation*}
+The binary liquid expansion of the free-energy model is presented in Appendix {\color{red} Put appendix here.
+Probably there is a special command like appendix.}
 
@@
 A.~Kuzmin wants to thank the Alberta Ingenuity Fund for their financial support.
+\section{Appendix A}
+{\color{red} Michal, you can add here binary Shan-Chen model. You have more information on it than me.}
+\section{Appendix B}
+\subsection{Binary liquid multiphase model}
+In this section we describe the binary liquid model. Binary model is based on
+the free energy model \cite{swift,landau}:
+\begin{equation}
+\mathfrak{F}=\int{J \mathrm{d}V}=\int{\left(c_s^2\rho\ln\rho+\frac{A}{2}\phi^2+\frac{B}{4}\phi^4 + \frac{k}{2}(\nabla \phi)^2 \right)\mathrm{d}V},
+\end{equation}
+where coefficients $A$ and $B$ are responsible for the phase separation,
+coefficient $k$ is responsible for the surface tension between phases and
+$\phi$ is the order parameter. Different phases are corresponding to
+extreme values of $\phi$ as $-1$ and $1$.
+The lattice Boltzmann model for the binary free-energy model is defined
+through two distribution sets:
+\begin{equation}
+\begin{aligned}
+&f_i^{eq}=w_i \rho \biggl(1+\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\biggr), i=0\div8\\
+&g_i^{eq}=w_i(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi^m \frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}), i\neq0 \\
+&g_0^{eq}=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
+\end{aligned}
+\end{equation}
+where $\Gamma$ is the mobility parameter, chemical potential
+$\mu=-A\phi+A\phi^3-k\Delta\phi$, $k$ is related to the surface
+tension. The macroscopic parameters are calculated as:
+\begin{equation}
+\begin{aligned}
+\rho=\sum_i{f_i}\\
+\rho u_{\alpha}=\sum_i{f_i c_{i\alpha}}\\
+\phi=\sum_i{g_i}
+\end{aligned}
+\end{equation}
+The equilibrium distribution functions lead to the macroscopic equations
+{\color{red} Give citation here}:
+\begin{equation}
+\begin{aligned}
+&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
+&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=\\
+&-\partial_{\alpha}P_{\alpha \beta} + \nu\partial_{\beta}\left(\partial_{\alpha}u_{\beta}+\partial_{\beta} u_{\alpha} + \frac{1}{3}\partial_{\gamma} u_{\gamma} \delta_{\alpha\beta}\right)\\
+&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=D\Delta \mu,
+\end{aligned}
+\label{binary:fluid:system}
+\end{equation}
+with the bulk pressure $p_0=\rho c_s^2-\frac{B}{2}\phi^2+\frac{3}{4}B \phi^2$.
+
+{\color{red} Michal, need your help here. Which parameters have you used for simulation. Please check above.}
+We take equilibrium distribution function for the binary-liquid gas model. In general, it can be presented in the following form:
 

### 2010-04-28 paper.tex
Fix appendix references.

@@
 \end{equation}
-The Shan-Chen model is one-component multiphase model. The multicomponent modification is presented in the Appendix {\color{red} put appendix reference here}.
+The Shan-Chen model is one-component multiphase model. The multicomponent modification is presented in the Appendix~\ref{sec:binary_shan_chen}.
 \subsection{Free-energy model}
@@
 We implement in terms of convergence,accuracy and performance this model with the Shan-Chen potential, mainly $p_0=\dfrac{\rho}{3}+G\dfrac{\psi^2(\rho)}{6}$.
-The binary liquid expansion of the free-energy model is presented in Appendix {\color{red} Put appendix here.
-Probably there is a special command like appendix.}
+The binary liquid expansion of the free-energy model is presented in Appendix~\ref{sec:binary_free_energy}
 
@@
 A.~Kuzmin wants to thank the Alberta Ingenuity Fund for their financial support.
-\section{Appendix A}
+
+\appendix
+\section{Binary fluid Shan-Chen model}
+\label{sec:binary_shan_chen}
 {\color{red} Michal, you can add here binary Shan-Chen model. You have more information on it than me.}
-\section{Appendix B}
-\subsection{Binary liquid multiphase model}
+\section{Binary liquid multiphase model}
+\label{sec:binary_free_energy}
 In this section we describe the binary liquid model. Binary model is based on

### 2010-04-28 paper.tex
Add Q_iab definition, rewrap a few paragraphs.

@@
 
-The lattice Boltzmann equation operates on the rectangular grid representing the physical domain.
-It utilizes the probability distribution functions containing the information about the macroscopic variables, such as density, momentum.
-LBE consists from two parts: local collision and propagation operated from one node to another in the certain direction specified by the velocity set.
-The lattice Boltzmann implementation in a standard view representing the collision and propagation steps:
+The lattice Boltzmann equation operates on the rectangular grid representing the physical domain. It utilizes the
+probability distribution functions containing the information about the macroscopic variables, such as density,
+momentum. LBE consists from two parts: local collision and propagation operated from one node to another in the certain
+direction specified by the velocity set.  The lattice Boltzmann implementation in a standard view representing the
+collision and propagation steps:
 \begin{equation}
@@
 \end{equation}
-where for D2Q9 $w_i=\{\dfrac{4}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{36},\dfrac{1}{36},\dfrac{1}{36},\dfrac{1}{36}\}$ and $Q_{i\alpha\beta}=$ with the sound speed parameter $c_s^2=\dfrac{1}{3}$.
-The macroscopic variables are restored on every time step through the probability function population:
+where for D2Q9
+$w_i=\{\dfrac{4}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{36},\dfrac{1}{36},\dfrac{1}{36},\dfrac{1}{36}\}$
+and $Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with the sound speed parameter $c_s^2=\dfrac{1}{3}$. The macroscopic variables are restored on
+every time step through the probability function population:
 \begin{equation*}
@@
 \end{equation*}
-The multiphase models can be simulated either through the force population and/or special equilibrium function.
-The Shan-Chen model utilizes the force population, but the free-energy model is simulated through the special equilibrium model.
-Both models are described in the next sections.
+The multiphase models can be simulated either through the force population and/or special equilibrium function. The
+Shan-Chen model utilizes the force population, but the free-energy model is simulated through the special equilibrium
+model. Both models are described in the next sections.
 

### 2010-04-28 paper.tex
Minor typographical fixes.

--- 2010-04-28 paper.tex
@@
 \end{equation}
-where for D2Q9 $w_i=\{\dfrac{4}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{36},\dfrac{1}{36},\dfrac{1}{36},\dfrac{1}{36}\}$ and $Q_{i\alpha\beta}=$ with the sound speed parameter $c_s^2=\dfrac{1}{3}$.
-The macroscopic variables are restored on every time step through the probability function population:
+where for D2Q9
+$w_i=\left\{\dfrac{4}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{36},\dfrac{1}{36},\dfrac{1}{36},\dfrac{1}{36}\right\}$
+and $Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with the sound speed parameter $c_s^2=\dfrac{1}{3}$. The macroscopic variables are restored on
+every time step through the probability function population:
 \begin{equation*}
@@
 \end{equation}
-where $w_{ixx}=\{\dfrac{1}{3},\dfrac{1}{3},-\dfrac{1}{6},-\dfrac{1}{6},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24}\}$, $w_{iyy}=\{\dfrac{1}{3},\dfrac{1}{3},-\dfrac{1}{6},-\dfrac{1}{6},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24}\}$ and $w_{ixy}=\{0,0,0,0,\dfrac{1}{4},\dfrac{1}{4},-\dfrac{1}{4},-\dfrac{1}{4}\}$.
-The laplacian and the gradient are taken by the numerical stencils:
+where $w_{ixx}=\left\{\dfrac{1}{3},\dfrac{1}{3},-\dfrac{1}{6},-\dfrac{1}{6},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24}\right\}$,
+$w_{iyy}=\left\{\dfrac{1}{3},\dfrac{1}{3},-\dfrac{1}{6},-\dfrac{1}{6},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24}\right\}$
+and $w_{ixy}=\left\{0,0,0,0,\dfrac{1}{4},\dfrac{1}{4},-\dfrac{1}{4},-\dfrac{1}{4}\right\}$. The laplacian and the gradient are taken by the numerical stencils:
 \begin{equation}

### 2010-04-28 codes/JuliaShanChen/cpu.cpp
Change the wxx, wxy and wyy coefficients so that they're in the correct (?) order.
This appears to prevent the simulation from blowing up, but it still needs to
be confirmed that the code is correct now.

@@ -72,7 +72,28 @@ int main(int argc, char** argv)
 	int i;
 	float dense,v1,v2;
 	
-	
+	int cx[]={0,1,0,-1,0,1,-1,-1,1};
+	int cy[]={0,0,1,0,-1,1,1,-1,-1};
+	float weights[]={4.0/9.0,1.0/9.0,1.0/9.0,1.0/9.0,1.0/9.0,1.0/36.0,1.0/36.0,1.0/36.0,1.0/36.0};
+//	float wxx[]={0.0,1.0/3.0,1.0/3.0,-1.0/6.0,-1.0/6.0,-1.0/24.0,-1.0/24.0,-1.0/24.0,-1.0/24.0};
+//	float wyy[]={0.0,-1.0/6.0,-1.0/6.0,1.0/3.0,1.0/3.0,-1.0/24.0,-1.0/24.0,-1.0/24.0,-1.0/24.0};
+//	float wxy[]={0.0,0.0,0.0,0.0,0.0,1.0/4.0,1.0/4.0,-1.0/4.0,-1.0/4.0};
+
+	float wxx[] = {0.0, 1.0/3.0, -1.0/6.0, 1.0/3.0, -1.0/6.0, -1.0/24.0, -1.0/24.0, -1.0/24.0, -1.0/24.0};
+	float wyy[] = {0.0, -1.0/6.0, 1.0/3.0, -1.0/6.0, 1.0/3.0, -1.0/24.0, -1.0/24.0, -1.0/24.0, -1.0/24.0};
+	float wxy[] = {0.0, 0.0, 0.0, 0.0, 0.0, 1.0/4.0, -1.0/4.0, 1.0/4.0, -1.0/4.0};
+
+	float gradstencilx[9]={0.0,4.0/12.0,0.0,-4.0/12.0,0.0,
+			          1.0/12.0,-1.0/12.0,-1.0/12.0,1.0/12.0};
+
+	float gradstencily[9]={0.0,0.0,4.0/12.0,0.0,-4.0/12.0,
+			          1.0/12.0,1.0/12.0,-1.0/12.0,-1.0/12.0};
+
+
+	float laplacestencil[9]={-20.0/6.0,4.0/6.0,4.0/6.0,4.0/6.0,4.0/6.0,
+					   1.0/6.0,1.0/6.0,1.0/6.0,1.0/6.0};
+
+
 	
 	for (i=0; i<nx*ny; i++)
 	{
[…]

### 2010-05-02 paper.tex
Fix equilibrium expression for the single fluid free energy model.
Fix wxx, wyy and wxy coefficients in the paper.
Added the Binary liquid test description. Need to calculate profiles. Battery runs out..

@@
 \begin{equation}
+\label{shan:chen:equation:state}
 P=c_s^2 \rho +\frac{G}{6} \psi(\rho)^2.
@@
 \subsection{Free-energy model}
-The free-energy or pressure model \cite{swift} allows the incorporation of the any equation of state to the lattice Boltzmann equation.
-In comparison with the Shan-Chen model, the free-energy model is simulated through the special choice of the equilibrium distribution function.
-The lattice Boltzmann scheme is implemented in a standard form (\ref{standard:implementation}), but the equilibrium function is defined as:
-\begin{equation}
+The free-energy or pressure model \cite{swift} allows the incorporation of the
+any equation of state to the lattice Boltzmann equation. In comparison with the
+Shan-Chen model, the free-energy model is simulated through the special choice
+of the equilibrium distribution function. The lattice Boltzmann scheme is
+implemented in a standard form (\ref{standard:implementation}), but the
+equilibrium function is defined as: \begin{equation}
 \begin{aligned}
-f_i^{eq}=&w_i \biggl(p_0(\rho)-3 k\rho\Delta\rho+\rho \frac{c_{i\alpha}u_{\alpha}}{3}+\rho \frac{9 Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2}\biggr)\\
-&+k (w_{ixx} \partial_x \rho \partial_x \rho + w_{ixy} \partial_x \rho \partial_y \rho + w_{iyy} \partial_y \rho \partial_y \rho),
+f_i^{eq}=&w_i \biggl(p_0(\rho)-3 k\rho\Delta\rho+\rho \frac{c_{i\alpha}u_{\alpha}}{c_s^2}+\rho \frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\biggr)\\
+&+k (w_{ixx} \partial_x \rho \partial_x \rho + w_{ixy} \partial_x \rho \partial_y \rho + w_{iyy} \partial_y \rho \partial_y \rho), \quad i > 0
 \end{aligned}
 \end{equation}
-where $w_{ixx}=\left\{\dfrac{1}{3},\dfrac{1}{3},-\dfrac{1}{6},-\dfrac{1}{6},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24}\right\}$,
-$w_{iyy}=\left\{\dfrac{1}{3},\dfrac{1}{3},-\dfrac{1}{6},-\dfrac{1}{6},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24}\right\}$
-and $w_{ixy}=\left\{0,0,0,0,\dfrac{1}{4},\dfrac{1}{4},-\dfrac{1}{4},-\dfrac{1}{4}\right\}$. The laplacian and the gradient are taken by the numerical stencils:
+where $w_{ixx}=\left\{\dfrac{1}{3},-\dfrac{1}{6},\dfrac{1}{3},-\dfrac{1}{6},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24}\right\}$,
+$w_{iyy}=\left\{-\dfrac{1}{6},\dfrac{1}{3},-\dfrac{1}{6},\dfrac{1}{3}, -\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24}\right\}$
+and $w_{ixy}=\left\{0,0,0,0,\dfrac{1}{4},-\dfrac{1}{4},\dfrac{1}{4},-\dfrac{1}{4}\right\}$. The laplacian and the gradient are taken by the numerical stencils:
 \begin{equation}
@@
 \label{sec:benchmark}
+A few benchmarks were made for comparison purposes between the Shan-Chen and the free-energy model:
+\begin{itemize}
+\item The first one compares accuracy and performance of calculations of the plane interface between gas and liquid phases.
+We included the Shan-Chen equation of state (\ref{shan:chen:equation:state}) to the free-energy model for the purpose of comparison.
+The domain {\color{red} Michal, please add the domain size} was initialized with the equilibrium densities calculated from Maxwellian area rule for $G=-5.5$.
+Half domain was initialized with the liquid density $\rho_l=...$, the another half domain was initialized with the gas density $\rho_g=...$.
+While the prescription of $G$ is enough for the Shan-Chen model to run simulations, the free-energy model operates with the additional parameter $k$ controlling the surface tension.
+We take values of $k$ {\color{red} Michal, please put a value here.} to guarantee the closeness of calculated profiles to the Shan-Chen model profiles \cite{sbragaglia-contact, kuzmin-thesis}.
 
+While the profiles are not compared since the Shan-Chen model has the density dependent surface tension coefficient \cite{shan-chen-free}, the models are compared in terms of accuracy to reproduce the bulk densities and the computer performance.
+The following outlines the findings: {\color{red} Michal, please add the benchmarks here}.
+
+\item The binary-liquid models can be compared for the two-phase Poiseuille flow.
+The flow is driven by pressure gradient implemented as the force, $\frac{\mathrm{d}P}{\mathrm{d}x}$.
+The walls are implemented as a half-way bounce-back.
+It is known that BGK fails to reproduce the exact location of the wall.
+Instead of BGK we used the D2Q9 MRT model \cite{dhumieres-first,dhumieres}.
+While BGK operates with one parameter $\tau$, the MRT model operates with the number of eigenvalues, i.e.
+$s_{\rho},s_e,s_{\epsilon},s_{j_x},s_{q_x},s_{j_y},s_{q_y},s_{p_{xx}},s_{p_{xy}}$, giving better stability and proper the Navier-Stokes equation restoration by varying the free model eigenvalues \cite{dhumieres}.
+For example for the half-way bounce-back walls representation for the Poiseuille flow the location of the wall is located exactly in the middle between nodes if the energy fluxes eigenvalues are defined as \cite{ginzburg-multireflection}:
+\begin{equation}
+\label{poiseuille:eigenvalues}
+s_{q_{x,y}}=8\frac{2-\omega}{8-\omega},
+\end{equation}
+where $\omega$ states for the $s_{p_{xx}}$ and $s_{p_{xy}}$ related to the bulk and kinematic viscosities.
+
+The domain {\color{red} Please put size here} is utilized with the two different densities {\color{red} What are they?}.
+The nodes with $y=0$ and $y=Ny$ represent the half-way bounce-back nodes.
+The MRT energy-flux eigenvalues are defined through the equation \ref{poiseuille:eigenvalues}, while viscosity related eigenvalues are defined as $s_{p_{xx}}=s_{p_{xy}}=\omega=1$.
+The velocity profile should obey the following formulae \cite{ginzburg-discontinious}:
+\begin{equation}
+\text{Need to calculate profiles. Michal, if you already have it - please add.}
+\end{equation}
+
+We compare the models in terms of the accuracy for velocity profiles for different density and viscosity ratios and the calculation performance.
+The following summarizes our findings: {\color{red} Michal, please add here benchmarks}.
+\end{itemize}
 \section{Conclusion}

### 2010-05-25 paper.tex
Rewrap a number of paragraphs.

@@
 \begin{itemize}
-\item The first one compares accuracy and performance of calculations of the plane interface between gas and liquid phases.
-We included the Shan-Chen equation of state (\ref{shan:chen:equation:state}) to the free-energy model for the purpose of comparison.
-The domain {\color{red} Michal, please add the domain size} was initialized with the equilibrium densities calculated from Maxwellian area rule for $G=-5.5$.
-Half domain was initialized with the liquid density $\rho_l=...$, the another half domain was initialized with the gas density $\rho_g=...$.
-While the prescription of $G$ is enough for the Shan-Chen model to run simulations, the free-energy model operates with the additional parameter $k$ controlling the surface tension.
-We take values of $k$ {\color{red} Michal, please put a value here.} to guarantee the closeness of calculated profiles to the Shan-Chen model profiles \cite{sbragaglia-contact, kuzmin-thesis}.
+\item The first one compares accuracy and performance of calculations of the
+	plane interface between gas and liquid phases.  We included the Shan-Chen
+	equation of state (\ref{shan:chen:equation:state}) to the free-energy model
+	for the purpose of comparison. The domain {\color{red} Michal, please add
+	the domain size} was initialized with the equilibrium densities calculated
+	from Maxwellian area rule for $G=-5.5$. Half domain was initialized with
+	the liquid density $\rho_l=...$, the another half domain was initialized
+	with the gas density $\rho_g=...$. While the prescription of $G$ is enough
+	for the Shan-Chen model to run simulations, the free-energy model operates
+	with the additional parameter $k$ controlling the surface tension. We take
+	values of $k$ {\color{red} Michal, please put a value here.} to guarantee
+	the closeness of calculated profiles to the Shan-Chen model profiles
+	\cite{sbragaglia-contact, kuzmin-thesis}.
 
-While the profiles are not compared since the Shan-Chen model has the density dependent surface tension coefficient \cite{shan-chen-free}, the models are compared in terms of accuracy to reproduce the bulk densities and the computer performance.
-The following outlines the findings: {\color{red} Michal, please add the benchmarks here}.
+While the profiles are not compared since the Shan-Chen model has the density
+dependent surface tension coefficient \cite{shan-chen-free}, the models are
+compared in terms of accuracy to reproduce the bulk densities and the computer
+performance. The following outlines the findings: {\color{red} Michal, please
+add the benchmarks here}.
 
-\item The binary-liquid models can be compared for the two-phase Poiseuille flow.
-The flow is driven by pressure gradient implemented as the force, $\frac{\mathrm{d}P}{\mathrm{d}x}$.
-The walls are implemented as a half-way bounce-back.
-It is known that BGK fails to reproduce the exact location of the wall.
-Instead of BGK we used the D2Q9 MRT model \cite{dhumieres-first,dhumieres}.
-While BGK operates with one parameter $\tau$, the MRT model operates with the number of eigenvalues, i.e.
-$s_{\rho},s_e,s_{\epsilon},s_{j_x},s_{q_x},s_{j_y},s_{q_y},s_{p_{xx}},s_{p_{xy}}$, giving better stability and proper the Navier-Stokes equation restoration by varying the free model eigenvalues \cite{dhumieres}.
-For example for the half-way bounce-back walls representation for the Poiseuille flow the location of the wall is located exactly in the middle between nodes if the energy fluxes eigenvalues are defined as \cite{ginzburg-multireflection}:
+\item The binary-liquid models can be compared for the two-phase Poiseuille
+	flow. The flow is driven by pressure gradient implemented as the force,
+	$\frac{\mathrm{d}P}{\mathrm{d}x}$. The walls are implemented as a half-way
+	bounce-back. It is known that BGK fails to reproduce the exact location of
+	the wall. Instead of BGK we used the D2Q9 MRT model
+	\cite{dhumieres-first,dhumieres}. While BGK operates with one parameter
+	$\tau$, the MRT model operates with a number of eigenvalues, i.e.
+	$s_{\rho},s_e,s_{\epsilon},s_{j_x},s_{q_x},s_{j_y},s_{q_y},s_{p_{xx}},s_{p_{xy}}$,
+	giving better stability and proper the Navier-Stokes equation restoration
+	by varying the free model eigenvalues \cite{dhumieres}.  For example for
+	the half-way bounce-back walls representation for the Poiseuille flow the
+	location of the wall is located exactly in the middle between nodes if the
+	energy fluxes eigenvalues are defined as \cite{ginzburg-multireflection}:
 \begin{equation}
@@
 
-The domain {\color{red} Please put size here} is utilized with the two different densities {\color{red} What are they?}.
-The nodes with $y=0$ and $y=Ny$ represent the half-way bounce-back nodes.
-The MRT energy-flux eigenvalues are defined through the equation \ref{poiseuille:eigenvalues}, while viscosity related eigenvalues are defined as $s_{p_{xx}}=s_{p_{xy}}=\omega=1$.
-The velocity profile should obey the following formulae \cite{ginzburg-discontinious}:
+The domain {\color{red} Please put size here} is utilized with the two
+different densities {\color{red} What are they?}. The nodes with $y=0$ and
+$y=Ny$ represent the half-way bounce-back nodes. The MRT energy-flux
+eigenvalues are defined through the equation \ref{poiseuille:eigenvalues},
+while viscosity related eigenvalues are defined as
+$s_{p_{xx}}=s_{p_{xy}}=\omega=1$. The velocity profile should obey the
+following formulae \cite{ginzburg-discontinious}:
 \begin{equation}
@@
 
-We compare the models in terms of the accuracy for velocity profiles for different density and viscosity ratios and the calculation performance.
-The following summarizes our findings: {\color{red} Michal, please add here benchmarks}.
+We compare the models in terms of the accuracy for velocity profiles for
+different density and viscosity ratios and the calculation performance. The
+following summarizes our findings: {\color{red} Michal, please add here
+benchmarks}.
+
 \end{itemize}

### 2010-05-25 sources/twophase_poiseuille.tex
Add a derivation of the Poiseuille profile in the 2D, two-phase case.

\documentclass[a4paper]{report}
\usepackage{amsmath}
\usepackage{fullpage}
\begin{document}

We are solving the Navier-Stokes equations in the 2D Cartesian coordinate system, assuming the system is driven by an
external acceleration field $g$ (in the Y direction):
\begin{align}
	\rho_i \left( \frac{\partial v_{ix}}{\partial t} + v_{ix} \frac{\partial v_{ix}}{\partial x} + v_{iy} \frac{\partial v_{ix}}{\partial y} \right) &= \mu_i
		\left( \frac{\partial^2 v_{ix}}{\partial x ^2 } + \frac{\partial^2 v_{ix}}{\partial y^2} \right) \\
	\rho_i \left( \frac{\partial v_{iy}}{\partial t} + v_{ix} \frac{\partial v_{iy}}{\partial x} + v_{iy} \frac{\partial v_{iy}}{\partial y} \right) &= \mu_i
		\left( \frac{\partial^2 v_{iy}}{\partial x ^2 } + \frac{\partial^2 v_{iy}}{\partial y^2} \right) + g \rho_i
\end{align}
We assume the flow to be stationary, so the time derivatives are 0.  We also take $v_{ix} = 0$, so the first equation
does not give us any further information about the system, while the second equation reduces to:
\begin{equation}
	- \frac{g \rho_i}{\mu_i} = \frac{\partial^2 v_{iy}}{\partial x^2}
\end{equation}
if we also asuume $\partial v / \partial y = 0$ (the flow is fully developed).  The general solution is:
\begin{equation}
	v_{iy} = \frac{g \rho_i}{2 \mu_i} x^2 + C_i x + D_i
	\label{eq:gensol}
\end{equation}
If we assume the phase interface to be located at $x = 0$, the continuity of $v$ requires $D_1 = D_2 = D$ and continuity of $\mu_i \partial_x v_i$ requires $\mu_1 C_1 = \mu_2 C_2$.

The flow happen in a channel of radius $R$, so no-slip boundary conditions are enforced at $x = -R$ and $x = R$:
\begin{align}
	v_{1y}(-R) &= \frac{g \rho_1}{2 \mu_1} R^2 - C_1 R + D = 0\\
	v_{2y}(R) &= \frac{g \rho_2}{2 \mu_2} R^2 + \frac{\mu_1}{\mu_2} C_1 R + D = 0
	\label{eq:vel}
\end{align}
Multiplying the first equation by $\mu_1 / \mu_2$ and adding the result to the second equation, we get:
\begin{equation}
	- \frac{g}{2 \mu_2} (\rho_1 + \rho_2) R^2 = D \left( 1 + \frac{\mu_1}{\mu_2} \right)
\end{equation}
and:
\begin{equation}
	D = -\frac{g (\rho_1 + \rho_2)}{ 2 (\mu_1 + \mu_2)} R^2
\end{equation}
Inserting that into Eq.~\ref{eq:vel}:
\begin{equation}
	R \left(\frac{g \rho_1}{2 \mu_1} -\frac{g (\rho_1 + \rho_2)}{ 2 (\mu_1 + \mu_2)} \right) = C_1
\end{equation}
which can be simplified to:
\begin{equation}
	C_1 = \frac{g R}{2} \frac{\mu_2 \rho_1 - \mu_1 \rho_2 }{\mu_1 (\mu_1 + \mu_2)}
\end{equation}

If we take the viscosities to be the same $\mu_1 = \mu_2 = \mu$, we get:
\begin{align}
	D &= -\frac{g (\rho_1 + \rho_2)}{4 \mu} R^2 \\
	C_2 = C_1 &= \frac{g (\rho_1 - \rho_2)}{4 \mu} R
\end{align}
The velocity profiles then read:
\begin{align}
	v_{1y}(x) &= \frac{g}{4 \mu} \left(2 x^2 - (\rho_1 - \rho_2) R x - (\rho_1 + \rho_2) R^2\right),\qquad x < 0\\
	v_{2y}(x) &= \frac{g}{4 \mu} \left(2 x^2 + (\rho_1 - \rho_2) R x - (\rho_1 + \rho_2) R^2\right),\qquad x > 0 
\end{align}

\end{document}

### 2010-05-25 sources/twophase_poiseuille.tex
Fix sings in the 2phase Poiseuille derivation.

@@
 We are solving the Navier-Stokes equations in the 2D Cartesian coordinate system, assuming the system is driven by an
-external acceleration field $g$ (in the Y direction):
+external acceleration field $-g$ (in the Y direction):
 \begin{align}
@@
 	\rho_i \left( \frac{\partial v_{iy}}{\partial t} + v_{ix} \frac{\partial v_{iy}}{\partial x} + v_{iy} \frac{\partial v_{iy}}{\partial y} \right) &= \mu_i
-		\left( \frac{\partial^2 v_{iy}}{\partial x ^2 } + \frac{\partial^2 v_{iy}}{\partial y^2} \right) + g \rho_i
+		\left( \frac{\partial^2 v_{iy}}{\partial x ^2 } + \frac{\partial^2 v_{iy}}{\partial y^2} \right) - g \rho_i
 \end{align}
@@
 \begin{equation}
-	- \frac{g \rho_i}{\mu_i} = \frac{\partial^2 v_{iy}}{\partial x^2}
+	\frac{g \rho_i}{\mu_i} = \frac{\partial^2 v_{iy}}{\partial x^2}
 \end{equation}
